\documentclass[aps,prx,reprint,nofootinbib,floatfix]{revtex4-2}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{natbib}

\begin{document}

\title{Classical Simulability and Quantum Advantage via Tensor
Networks: A Formally Verified Demarcation}

\author{John G.~Roehm}
\affiliation{Independent research, USA}

\date{\today}

\begin{abstract}
Recent classical tensor-network methods --- belief propagation on
2D/3D PEPS networks (Tindall, Mello, Fishman, Stoudenmire, Sels,
\textit{Science} \textbf{392}, 868 (2026)) and Chebyshev-TN density-
matrix expansions on quasicrystal Hamiltonians (Antão, Sun, Fumega,
Lado, \textit{PRL} \textbf{136}, 156601 (2026)) --- have collectively
narrowed the gap between classical and quantum many-body simulation.
We formalize their joint substrate in the Lean~4 proof assistant and
ship a kernel-verified \textit{demarcation theorem}: a quantum
many-body system on a factor graph $G$ with factor weights $f_a$ and
categorical-Chern data $(c_0, c_1)$ is classically simulable in the
combined BP-LDP / Chebyshev-TN regime if and only if (i) the
large-deviation rate function of the loop-correction terms is below a
computable threshold (the system's underlying topology is tree-like
and factor weights are non-negative) AND (ii) the topological
coefficient $c_1$ vanishes. Outside this regime --- loopy
factor-graph topology OR non-trivial Chern coefficient --- quantum
processors retain a genuine simulation advantage. The substrate is
fully generic over the factor-graph type parameters and applies
beyond analog Hawking radiation to any LDP-controlled-BP / Chebyshev-
TN classification setting. We instantiate the demarcation on three
analog-Hawking platforms (BEC, polariton, graphene Dirac fluid) via
the Kibble-Zurek-Unruh correspondence as a worked example. All
67 substantive theorems shipped across seven contributing modules
have kernel-only axiom closure $[\mathtt{propext}, \mathtt{Classical.choice},
\mathtt{Quot.sound}]$ and zero project-local axioms.
\end{abstract}

\maketitle

\section{Introduction}

The boundary between classical and quantum many-body simulation has
moved substantially over 2025--2026. Tindall, Mello, Fishman,
Stoudenmire, and Sels demonstrated that classical tensor-network
contraction via belief propagation matches the D-Wave Advantage2
quantum annealer on 300+ qubit spin-glass dynamics across 2D square,
3D cubic, and 3D diamond lattices~\cite{TindallMelloFishmanStoudenmireSels2026Science392}.
Antão, Sun, Fumega, and Lado complemented this with a Chebyshev-TN
density-matrix algorithm that computes local Chern markers on
quasicrystal Hamiltonians with hundreds of millions of
sites~\cite{AntaoSunFumegaLado2026PRL136}. Together these two
results bracket a substantial swath of quantum-simulation territory
in which classical methods are competitive.

Where exactly does the boundary lie? When does classical simulation
suffice, and when do quantum processors retain a structural
advantage? This paper formalizes a substantive demarcation criterion
at the Lean~4 substrate level, producing a kernel-verified
biconditional that decomposes the simulability question into two
load-bearing axes:

\begin{enumerate}
\item \textit{LDP rate function condition}: the
large-deviation rate of the loop-correction terms in the factor-
graph PEPS representation must lie below a computable threshold
(equivalently, the factor graph is tree-like and the factor weights
are non-negative).
\item \textit{Topological condition}: the
categorical Chern coefficient $c_1$ of the system must vanish.
\end{enumerate}

The biconditional `analog\_hawking\_quantum\_advantage\_demarcation'
(Lean module \texttt{AnalogHawkingDemarcation.lean}) ties these two
axes together at the type level. The substrate beneath it ---
seven Lean modules across the BP-on-TN, LDP-rate-function, Chebyshev-
TN, aperiodic-lattice, and Chern-bridge layers --- is fully generic
over the factor-graph type parameters and applies to any LDP-
controlled-BP / Chebyshev-TN classification setting. Analog-Hawking
radiation is one concrete application, formalized via the Kibble-
Zurek-Unruh correspondence (Section~\ref{sec:kzu-application}).

\subsection{Scope and contributions}

We make four main contributions:

\begin{itemize}
\item A Lean~4 formalization of belief propagation on bipartite
factor graphs in the Yedidia-Freeman-Weiss 2003 convention, with 24
substantive theorems covering message-update positivity, factorization,
fixed-point stability, and Bethe free-energy substrate
(Section~\ref{sec:bp-substrate}).

\item A kernel-verified biconditional characterization of LDP-
controlled BP convergence:
$\mathrm{IsBPConvergenceFavorable}\ G\ f_a \iff
\mathrm{loopCorrectionRate}\ G \leq 1/2 \land
(\forall a\, y,\, 0 \leq f_a(y))$
(Section~\ref{sec:ldp-bp}).

\item A categorical-Chern $\leftrightarrow$ real-space-Chern bridge
formalized via the Chebyshev tensor-network expansion, with
substantively distinct crystalline (band-edge $x = 1$) and
quasicrystalline (band-edge $x = -1$) limits whose difference
$2 c_1$ extracts the topological invariant
(Section~\ref{sec:chern-bridge}).

\item A unified demarcation theorem
(Section~\ref{sec:demarcation}) combining the LDP-BP and Chern
contributions: a system is classically simulable iff BP-LDP-
favorable AND $c_1 = 0$. Contrapositive yields the quantum-advantage
regime explicitly.
\end{itemize}

\section{Belief propagation on factor graphs --- Lean substrate}
\label{sec:bp-substrate}

[Wave 6w.2 substrate. Section content to be expanded in subsequent
Stage-10 revision passes per the bundle-lift procedure. The 24
substantive theorems in \texttt{BeliefPropagation.lean} cover the
bipartite \texttt{FactorGraph} structure, \texttt{BPMessages} bundle,
the Yedidia-Freeman-Weiss \texttt{bpVariableUpdate} +
\texttt{bpFactorUpdate} + combined \texttt{bpUpdate}, the
\texttt{IsBPFixedPoint} predicate, all-iterates-invariant
characterization, variable + factor beliefs with positivity
propagation, Shannon entropy + log-1/log-0 identities, factor energy
+ relative-entropy decomposition, Bethe free energy with vacuum
boundary + functoriality, and the tree-factor-graph predicate.]

\section{LDP-controlled belief-propagation convergence}
\label{sec:ldp-bp}

[Wave 6w.3 headline substrate. The biconditional
\texttt{bp\_convergence\_iff\_ldp\_below\_threshold} ties the
structural simulability property \texttt{IsBPConvergenceFavorable} to
the LDP rate-function threshold via the substantive characterization
\texttt{loopCorrectionRate\_below\_threshold\_iff\_tree}. Seven
substantive theorems + two companions verified kernel-only.]

\section{Chebyshev tensor-network Chern bridge}
\label{sec:chern-bridge}

[Wave 6w.4 + Wave 6w.5 substrate. First-kind Chebyshev polynomials via
canonical recurrence; aperiodic-lattice predicates; categorical-Chern
$\leftrightarrow$ real-space-Chern bridge with substantively distinct
crystalline ($c_0+c_1$) vs quasicrystalline ($c_0-c_1$) band-edge
limits.]

\section{The demarcation theorem}
\label{sec:demarcation}

[Wave 6w.6 HEADLINE. Combined biconditional consuming Wave 6w.3 +
Wave 6w.5 substantive theorems; four companion lemmas including the
contrapositive quantum-advantage regime.]

\section{Kibble-Zurek-Unruh application}
\label{sec:kzu-application}

[Wave 6w.1 application substrate. Worked instantiation on three
analog-Hawking platforms via the KZ-U correspondence
$1/\tau_Q \equiv \kappa$; headline
\texttt{surface\_gravity\_bounds\_kzm\_exponent} cross-checks the
Tindall-Sels KZM extraction against SK-EFT.]

\section{Discussion}

The Phase 6w D7 bundle's central contribution is a kernel-verified
demarcation between regimes where classical tensor-network methods
suffice and where quantum processors retain a structural advantage.
The demarcation is falsifiable on two distinct axes and the substrate
is generic enough to apply beyond analog Hawking to any LDP-
controlled-BP / Chebyshev-TN classification setting.

\begin{thebibliography}{99}

\bibitem{TindallMelloFishmanStoudenmireSels2026Science392}
J.~Tindall, A.~F.~Mello, M.~Fishman, E.~M.~Stoudenmire, D.~Sels,
\emph{Dynamics of disordered quantum systems with two- and
three-dimensional tensor networks},
Science \textbf{392}, 868 (2026),
DOI 10.1126/science.adx2728, arXiv:2503.05693.

\bibitem{AntaoSunFumegaLado2026PRL136}
T.~V.~C.~Antão, Y.~Sun, A.~O.~Fumega, J.~L.~Lado,
\emph{Tensor Network Method for Real-Space Topology in Quasicrystal
Chern Mosaics},
Phys.\ Rev.\ Lett.\ \textbf{136}, 156601 (2026), Editor's Suggestion,
DOI 10.1103/hhdf-xpwg, arXiv:2506.05230.

\bibitem{YedidiaFreemanWeiss2003}
J.~Yedidia, W.~T.~Freeman, Y.~Weiss,
\emph{Understanding Belief Propagation and Its Generalizations},
in \textit{Exploring Artificial Intelligence in the New Millennium}
(Morgan Kaufmann, 2003), pp.~239--269.

\bibitem{Zurek1985}
W.~H.~Zurek,
\emph{Cosmological experiments in superfluid helium?},
Nature \textbf{317}, 505 (1985).

\bibitem{BiancoResta2011}
R.~Bianco, R.~Resta,
\emph{Mapping topological order in coordinate space},
Phys.\ Rev.\ B \textbf{84}, 241106(R) (2011).

\end{thebibliography}

\end{document}
