\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{array}
\usepackage[margin=1in]{geometry}

\lstset{basicstyle=\ttfamily\small, breaklines=true, frame=single}

%% Bundle target: I1 (tier 3, journal CPC | Phys. Rep.)
%% Schema: docs/BUNDLE_DIRECTORY_SCHEMA.md
%% Source manifest: papers/I1/source_manifest.md
%% Append log: papers/I1/append_log.json
%% Change log: papers/I1/change_log.md
%% This bundle is a synthesis-driven new composition (Phase 7a sub-wave
%% 7a.2). It draws on paper15_methodology, paper44_riemannian_connection,
%% Phase 6f W1-W6 algebraic-GR substrate rosters, and Phase 6g W5
%% structural-Prop scoping pattern as substrate; the prose is authored
%% fresh.
\input{../../docs/counts.tex}

\title{Continuous Verification for Computational Physics Research:\\
A Three-Layer Pipeline with Worked Cases from Schwinger--Keldysh
Effective Field Theory, Algebraic General Relativity, and
Topological Quantum Field Theory}

\author{John Roehm}
\date{Draft --- Phase 7a, May 2026}

\begin{document}

\maketitle

\begin{abstract}
Computational physics research lacks a structural guarantee that
published numerical claims match the underlying computation. We
describe a three-layer continuous-verification pipeline integrating
Python numerics, Lean~4 formal proofs, and the Aristotle automated
theorem prover, organized into a fourteen-stage workflow with
fourteen invariants. Applied across a research program spanning
dissipative effective field theory, analog Hawking radiation,
emergent gravity through black-hole thermodynamics, anomaly
constraints on Standard Model particle content, dark-sector substrate
constraints, and topological quantum computation foundations, the
pipeline has produced \totaltheorems{} machine-checked theorems
across \leanmodules{} Lean modules with \sorrycount{} remaining gaps
(\sorrypercent\%) plus \aristotleproved{} theorems closed by
automated theorem proving. We document three new pipeline stages
beyond the original twelve-stage version (interactive Lean
development via the language-server protocol, fresh-context
adversarial review, and meta-process quality improvement), three new
invariants (placeholder non-load-bearing, no heartbeat overrides in
proof bodies, primary-source citation cache), the publication-bundle
architecture that consolidates per-wave drafts into thirteen
publication targets, and the preemptive-strengthening discipline
that reduces theorem-statement audit cost from twelve cuts to zero
across recent waves. Three worked cases illustrate the methodology:
the Aristotle FirstOrderKMS counterexample (where automated proving
failed and the failure surfaced a substantive physics result), the
gap-solution-bounded conjecture disproof (where formalization-driven
counterexample search refuted a literature claim), and the
chirality-wall axiom decomposition (where structural axiom
classification turned a single 50-term proof into nine
individually-tractable sublemmas). Two additional sidebar studies
document first-in-any-proof-assistant formalization claims at the
algebraic-general-relativity substrate level and a
Mathlib-dependency-fallback discipline pattern for theorems whose
substantive content has no upstream PDE infrastructure available.
The methodology is domain-agnostic.
\end{abstract}

%% ─────────────────────────────────────────────────────────────────
%% §1 Introduction
%% ─────────────────────────────────────────────────────────────────

\section{Introduction}
\label{sec:intro}

Numerical computation in physics research is increasingly
sophisticated and increasingly opaque. A typical large
computational-physics program now combines several thousand lines of
Python or Julia, dozens of Jupyter notebooks, and a paper-trail of
PDF drafts whose numerical content is the output of code paths that
the published prose does not name. The standard publication workflow
provides no machine-checkable connection between the prose and the
code: reviewers cannot, and do not, run the computation. Five years
after publication, when the original code has bit-rotted and the
authors have rotated, no one can reconstruct which version of which
formula produced which figure.

This is not a hypothetical concern. We have personally observed
each of the following failure modes during a single research
program: (i) a numerical claim in a draft that traced to a code
path that no longer existed; (ii) a ``formally verified'' claim
whose underlying Lean theorem was a placeholder of type
\texttt{True}; (iii) a citation whose arXiv identifier resolved to
an unrelated graph-neural-network paper; (iv) a parameter whose
value was inherited from a prior draft and had drifted from the
primary source it cited. None of these were detected by the
authors before adversarial review. All of them shipped to draft
state.

The pipeline we describe makes verification \emph{continuous}, in
the sense of integrated into the daily research workflow rather
than retrofitted before submission. It has fourteen stages and
fourteen invariants, integrates three orthogonal verification
layers, and produces machine-checkable artifacts at each stage. We
have applied it to a research program spanning \papercount{}
papers covering Schwinger--Keldysh dissipative effective field
theory, analog Hawking radiation in Bose--Einstein condensates and
graphene, emergent gravity through black-hole thermodynamics,
anomaly constraints on Standard Model particle content, dark-sector
substrate constraints from causal-set theory, entropic gravity, and
Jacobson thermodynamic emergence, and topological quantum
computation foundations. The pipeline has produced
\totaltheorems{} machine-checked theorems across \leanmodules{}
Lean modules, with \sorrycount{} remaining unproved gaps
(\sorrypercent\%) and \aristotleproved{} theorems closed by the
Aristotle automated theorem prover.

This paper extends and supersedes an earlier description of a
twelve-stage version of the pipeline (paper15 methodology draft,
April 2026). The version described here adds three stages beyond
the original twelve --- interactive Lean development via the
language-server protocol (Stage~3a), fresh-context adversarial
review (Stage~13), and meta-process quality improvement
(Stage~14) --- and three invariants (placeholder non-load-bearing,
no heartbeat overrides in proof bodies, primary-source citation
cache). It also documents the
publication-bundle architecture that consolidates per-wave drafts
into thirteen publication targets and the preemptive-strengthening
discipline that reduces theorem-statement audit cost from twelve
retroactive cuts (one wave) to zero (subsequent waves) by applying
strengthening rules prospectively.

The remainder of this paper is organized as follows.
Section~\ref{sec:architecture} describes the three-layer
verification architecture (Python numerics, Lean~4 formal proofs,
Aristotle automated proving). Sections~\ref{sec:case1},
\ref{sec:case2}, \ref{sec:case3} present three worked cases that
illustrate the pipeline catching errors that single-layer
verification cannot. Section~\ref{sec:pipeline} presents the
fourteen-stage workflow and its invariants in detail.
Section~\ref{sec:strengthening} documents the preemptive-strengthening
discipline.
Section~\ref{sec:provenance} describes sentence-level provenance and
the cross-paper consistency cluster mechanism that enables bundle
publication. Section~\ref{sec:three_layer_extended} returns to the
three-layer architecture in the context of automated theorem proving
integration. Section~\ref{sec:adversarial} describes the
fresh-context adversarial-reviewer pattern.
Section~\ref{sec:phase6f_substrate} presents a sidebar of
first-formalization-in-any-proof-assistant claims at the
algebraic-general-relativity substrate level
(Phase~6f Waves~1--6). Section~\ref{sec:riemannian_sidebar}
presents the Phase~6f Wave~7 catch-up Riemannian connection sidebar
and its bundle-level extension via the abstract-bundle Levi-Civita
construction (Phase~6f Wave~8 Sessions~1--3).
Section~\ref{sec:structural_prop_scoping} describes the
Mathlib-dependency-fallback structural-Prop scoping discipline
(Phase~6g Wave~5). Section~\ref{sec:lessons} discusses lessons
learned and outlines future pipeline extensions.

%% ─────────────────────────────────────────────────────────────────
%% §2 Three-layer verification architecture
%% ─────────────────────────────────────────────────────────────────

\section{The Three-Layer Verification Architecture}
\label{sec:architecture}

The pipeline integrates three orthogonal verification layers, each
addressing a failure mode that the others cannot.

\textbf{Layer~1: Python numerics for fast iteration.} A formula is
implemented in a canonical \texttt{formulas.py} module, with a
docstring containing (i) the mathematical formula in plain text,
(ii) the Lean theorem name that formalizes it, (iii) the Aristotle
run identifier (or ``manual'' or ``pending''), and (iv) a citation
to the published source for derived formulas. Domain-specific
modules import from \texttt{formulas.py} rather than reimplementing.
Python tests verify computational correctness, edge cases, and
physical bounds. Layer~1 catches computation errors but not
formalization errors.

\textbf{Layer~2: Lean~4 formal proofs for mathematical certainty.}
Each formula is stated as a Lean theorem in the strongest form
possible. Statements are not weakened for proof convenience; if
the strongest form is not provable, the proof architecture is
wrong (Section~\ref{sec:strengthening}). Manual proofs are written
for theorems within reach. Layer~2 catches formalization errors
but not definition errors --- a formally verified proof can still
be wrong if the definitions it operates on are wrong
(Section~\ref{sec:case3}).

\textbf{Layer~3: Aristotle automated theorem prover for sorry
gaps.} The entire Lean project is submitted with strategy hints
embedded in the docstrings of unproved theorems (the ``PROVIDED
SOLUTION'' convention). Aristotle uses tactic synthesis, term
search, and machine learning to find proofs. Results are
manually reviewed before integration --- no blind acceptance ---
to verify that the proof exercises the hypotheses, that
\texttt{lake build} passes after integration, and that no
heartbeat overrides have been introduced (Invariant~10).

The three layers interact in a discovery--formalization--automation
cycle. A typical cycle: a Python implementation of a new formula
exposes a numerical regularity; the regularity is stated as a Lean
theorem with sorry; the theorem is closed by Aristotle in a
priority batch; the closure feeds back into Python by setting the
Aristotle run identifier in the formula docstring, which is then
checked by the validation suite (Section~\ref{sec:pipeline}). The
critical-coupling theorem $G_c = 8\pi^2/(N_f \Lambda^2)$ for the
chirality-wall mechanism completed all three layers in a single
session: derived in Python, stated in Lean, closed by Aristotle in
run \texttt{79e07d55}, and checked by validation, all within
twenty-four hours.

The composition is asymmetric in a way that matters. Layer~1 alone
gives reproducibility but not correctness: notebooks can produce
the same wrong answer indefinitely. Layer~2 alone gives correctness
but not numerical scope: a formalization of an idealized model is
silent on the actual experimental parameters.
Layer~3 alone gives proof-search throughput but no verification of
what was proved: an automated proof can be vacuous if the
definitions it operates on are wrong (Section~\ref{sec:case3}). The
composition gives all three.

%% ─────────────────────────────────────────────────────────────────
%% §3 Worked case 1: FirstOrderKMS Aristotle counterexample
%% ─────────────────────────────────────────────────────────────────

\section{Worked Case~1: The FirstOrderKMS Aristotle Counterexample}
\label{sec:case1}

% TODO: substantive draft. Substrate to draw from:
%  - SK-EFT Hawking project memory `feedback_aristotle_negative_results.md`
%  - First-order Kubo-Martin-Schwinger relation in
%    `lean/SKEFTHawking/FirstOrder.lean` (the theorem that Aristotle
%    repeatedly failed to close)
%  - The user-author iteration that surfaced the failure mode
%  - The eventual physics resolution: the FirstOrderKMS at zeroth
%    order is genuinely zero --- not because Aristotle was wrong,
%    but because the symmetry-protected zero is the substantive
%    claim, and the formal pipeline's inability to discover it
%    automatically was diagnostic, not pathological
%
% Section thesis: when a Lean theorem resists automated proof
% repeatedly, the failure is information --- it surfaces either a
% wrong-statement issue (P-anti-pattern, Section 7) or a
% missing-physics issue. The FirstOrderKMS case is the latter: the
% absence of an automated proof was diagnostic of the symmetry
% protection that the corresponding paper's central claim depends on.

\emph{Section to be drafted in subsequent session.} This section
presents the case where Aristotle repeatedly returned
\texttt{OUT\_OF\_BUDGET} on what appeared to be a routine
first-order Kubo--Martin--Schwinger (KMS) relation, and the failure
turned out to be diagnostic: the KMS relation at zeroth order is
genuinely zero by symmetry protection, not by computational
accident. The pipeline's inability to discover the symmetry
automatically surfaced the substantive physics claim.

%% ─────────────────────────────────────────────────────────────────
%% §4 Worked case 2: gap-solution-bounded conjecture disproof
%% ─────────────────────────────────────────────────────────────────

\section{Worked Case~2: The Gap-Solution-Bounded Conjecture Disproof}
\label{sec:case2}

% TODO: substantive draft. Substrate to draw from:
%  - The literature claim that a gap solution must be bounded
%  - The formalization-driven counterexample search that refuted it
%  - The Lean module containing the explicit unbounded counterexample
%  - The PR (or commit-pinned counterexample reference) that
%    documents the formal disproof
%
% Section thesis: formal verification is not just for confirming
% known results; the systematic structure of a Lean formalization
% can drive counterexample search that informal pen-and-paper work
% does not. In this case, attempting to prove the conjecture
% surfaced the precise sufficient hypothesis that was missing,
% which then exposed an explicit counterexample.

\emph{Section to be drafted in subsequent session.} This section
presents a literature claim (gap solution must be bounded) that
formalization in Lean could not reproduce despite repeated attempts,
where the failure pattern surfaced a missing sufficient
hypothesis, and where filling that gap exposed an explicit
counterexample. The counterexample is commit-pinned in the project
repository.

%% ─────────────────────────────────────────────────────────────────
%% §5 Worked case 3: chirality-wall axiom decomposition
%% ─────────────────────────────────────────────────────────────────

\section{Worked Case~3: The Chirality-Wall Axiom Decomposition}
\label{sec:case3}

% TODO: substantive draft. Substrate to draw from:
%  - Chirality-wall mechanism in EW baryogenesis context
%    (paper33_ewbg_chirality_wall, Phase 6c W2)
%  - The single 50-term theorem that initially could not be closed
%  - The decomposition into nine sub-lemmas, each <=12 terms
%  - Aristotle closure of the nine sub-lemmas in a single batch
%  - The structural-axiom classification (eliminable vs hard vs
%    unknown) that informed the decomposition
%
% Section thesis: a recurring pattern is that a monolithic proof
% goal is intractable not because it is hard but because it is
% large. Decomposition into structural sub-lemmas, each of which is
% individually tractable, is the standard fix. The chirality-wall
% case decomposed nine-deep and Aristotle closed all nine
% sub-lemmas in a single submission.

\emph{Section to be drafted in subsequent session.} This section
presents a chirality-wall theorem in the context of electroweak
baryogenesis where a fifty-term monolithic proof goal was
intractable to both manual proof and Aristotle, but where a
nine-fold structural decomposition produced sublemmas each
individually closable, and where Aristotle closed all nine in a
single priority batch. The structural-axiom classification (each
sublemma's eliminability tier per AXIOM\_METADATA) informed the
decomposition order.

%% ─────────────────────────────────────────────────────────────────
%% §6 Fourteen-stage pipeline
%% ─────────────────────────────────────────────────────────────────

\section{The Fourteen-Stage Wave Execution Pipeline}
\label{sec:pipeline}

The pipeline is organized as fourteen sequential stages, each with
a gate that must pass before the next stage begins. Skipping
stages or reordering them produces rework and allows errors to
propagate downstream. Wave-level work --- adding new physics,
extending formalization, or producing experimental predictions
--- proceeds through all fourteen.

\begin{table}[h]
\centering
\small
\input{tables/table1_stages.tex}
\caption{Pipeline stages with their gates, autogenerated from
\texttt{WAVE\_EXECUTION\_PIPELINE.md}.}
\label{tab:stages}
\end{table}

The current pipeline extends the original twelve-stage version
(paper15 draft, April 2026) with three additional stages.
Stage~3a, interactive Lean development via the language-server
protocol, replaces the previous ``write tactic, run \texttt{lake
build}, parse error, iterate'' development loop (approximately
fifteen seconds per cycle) with a millisecond-latency interactive
loop using the \texttt{lean-lsp-mcp} tool integration. The
interactive loop reads the proof state at any line and column,
tests candidate tactic sequences without editing the file, and
returns ``no goals'' when the proof is complete. Empirically, the
interactive loop is roughly a thousand times faster for proof
construction than the build-iterate cycle.

Stage~13, fresh-context adversarial review, runs a Lean-aware
adversarial reviewer agent in a fresh context window against each
paper or bundle near submission readiness. The agent works through
eight finding classes (citation integrity, parameter provenance,
narrative grounding, cross-paper consistency, count freshness,
formal-verification-claim substantiation, assumption disclosure,
and review-state hygiene) and emits a structured markdown report.
Findings are auto-picked up as graph nodes; ``BLOCKER'' findings
flip the affected readiness gate to blocked. Stage~13 is meaningful
only on a codebase that already passes its own internal validation
(Stages~1--12); it backstops the failure classes that internal
validation cannot detect by construction.

Stage~14, meta-process quality improvement, scans the cumulative
adversarial-review findings for systemic patterns: failure classes
that recur across multiple papers, or findings that indicate a gap
in the pipeline itself. Such patterns are recorded in a quality-
improvement register with owners and target dates. Stage~14 is
advisory; it never blocks publication. Its purpose is feeding
pipeline improvements back into subsequent waves. The April~2026
external-adversarial-review round, which surfaced a thirteen-
dimensional problem space across the project, was the seed event
for the register.

Stages~1--12 are the substantive computational and formalization
work. Stage~1 establishes the canonical sources of truth: physical
constants in \texttt{constants.py}, formulas in
\texttt{formulas.py}, figure functions in \texttt{visualizations.py}.
Stage~2 implements formulas with provenance metadata. Stages~3--4
formalize and prove (interactive plus automated). Stage~5 verifies
the formal layer compiles cleanly. Stage~6 runs Python tests.
Stage~7 runs sixteen automated cross-layer validation checks
(Section~\ref{sec:provenance}). Stage~8 generates publication-
quality figures. Stage~9 runs a figure-reviewer agent. Stage~10
drafts the paper with full provenance for every claim. Stage~11
produces reproducible computational notebooks. Stage~12 syncs
documentation and runs counts regeneration. Each stage's gate must
pass before proceeding.

The pipeline's invariants must hold at all times, not just at stage
completion:

\begin{enumerate}
\item \textbf{Canonical formulas.} \texttt{formulas.py} is the only
place physics formulas live.

\item \textbf{Canonical constants.} \texttt{constants.py} is the only
place experimental parameters and the theorem registry live.

\item \textbf{Canonical visualizations.} \texttt{visualizations.py}
is the only place figure functions live.

\item \textbf{Every formula has a Lean theorem.} Every Lean theorem
has a proof (zero sorry target).

\item \textbf{Every computed quantity has bounds.} Physical bounds
are tested and enforced by validation.

\item \textbf{Every paper claim traces to computation.} Numerical
claims match formula output within 0.5\%.

\item \textbf{Narrative derives from data.} Feasibility claims need
computed support.

\item \textbf{Every parameter has verified provenance.} Each value
traces to a published source via the
\texttt{PARAMETER\_PROVENANCE} table.

\item \textbf{Placeholder theorems are non-load-bearing.} Theorems
proved as \texttt{True := trivial} are documentation markers
only. Tracked in a \texttt{PLACEHOLDER\_THEOREMS} registry; not
counted as substantive.

\item \textbf{No heartbeat overrides in proof bodies.} A proof body
that hits the heartbeat limit is evidence that the proof
architecture is wrong; the fix is sub-lemma decomposition, not
raising the budget.

\item \textbf{Every external bibitem has a primary-source cache
file.} Each non-in-prep entry in the citation registry carries a
path to a primary-source artifact under
\texttt{Lit-Search/Phase-X/primary-sources/}, structurally
preventing the hallucinated-citation failure mode.

\item \textbf{Provenance DOIs resolve to the registry.} Every DOI
mentioned in any parameter-provenance source must resolve to a
registry bibkey; every cited bibkey must exist in the registry.

\item \textbf{Stage~14 quality-improvement register is auto-
regenerated and manually curated, never wiped.} Closed items
are preserved verbatim across regenerations.

\item \textbf{Every paper-shaped output lifts into a publication
bundle.} New drafts, section additions, and notebook companions
identify their target bundle at Stage~1 and record it in the
draft-mapping document at Stage~12. No new stand-alone paper
drafts are created without explicit user authorization to add a
fourteenth-or-higher bundle target.

\end{enumerate}

Invariant~9 addresses a subtle inflation problem:
\placeholdertheorems{} of our \totaltheorems{} theorems are
\texttt{True := trivial} placeholders that serve as documentation
markers. These are tracked in a registry
(\texttt{PLACEHOLDER\_THEOREMS}) and excluded from substantive
theorem counts; the substantive count is \substantivetheorems{}.

Invariant~14 is the publication-bundle architecture, described in
Section~\ref{sec:provenance}.

%% ─────────────────────────────────────────────────────────────────
%% §7 Preemptive-strengthening discipline
%% ─────────────────────────────────────────────────────────────────

\section{The Preemptive-Strengthening Discipline}
\label{sec:strengthening}

The strengthening pass is an audit that runs at the end of each
wave, examining each new theorem statement for six recurring
anti-patterns. Originally retroactive --- the audit ran after
theorems were stated and proved, with retroactive cuts applied
where the audit identified a vacuous statement --- the pass has
been moved to the prospective phase: each theorem statement is
checked against the same six anti-patterns \emph{before} the
statement is written. The reduction in audit cost is
substantial: the wave that introduced the discipline (Phase~6c.3)
required twelve retroactive cuts; the next wave (Phase~6b.1)
required five; the wave after that (Phase~6m R6) required zero.

The six anti-patterns are:

\begin{itemize}

\item \textbf{P2 --- Bundle redundancy.} A theorem stating a
conjunction is redundant if dropping any conjunct produces a
statement of equal information content. The pattern occurs when
multiple conjuncts are algebraically equivalent: for instance,
$\mathrm{satisfiesAt}\ m\ \mathrm{WEP}\ \wedge\ \mathrm{satisfiesAt}\
m\ \mathrm{EEP}\ \wedge\ \mathrm{satisfiesAt}\ m\ \mathrm{SEP}$ when
$\mathrm{violationLevel}\ m = \mathrm{none}$ is one substantive
theorem dressed as three. The fix is to write the substantive form
($\mathrm{violationLevel}\ m = \mathrm{none}$) plus a separate
extraction lemma if downstream callers need the bundle shape.

\item \textbf{P3 --- Multiplication-only triviality.} A theorem
whose substantive content is exhausted by an obvious arithmetic
identity ($0.34 = 2 \times 0.17$, for instance) is structural, not
physical. The fix is to replace with a load-bearing comparison
that separates a conformant case from a violator
($0.34 < 1.0$).

\item \textbf{P4 --- Self-discharging axiom.} An axiom whose
statement reduces to truth via algebra alone is vacuous. The fix
is to either remove the axiom or write the substantive content
elsewhere.

\item \textbf{P5 --- Falsifier as restated hypothesis.} A
``falsifier'' theorem whose body restates its hypotheses without
exercising any computation is structural-tautology repackaging.
The fix is to ensure the falsifier exercises the hypothesis.

\item \textbf{P6 --- Cross-module bridge integrity.} If a
theorem's docstring references another module's theorem, the body
must \texttt{import} that module and \emph{call} the theorem ---
not merely list it as a documentation pointer. The audit pattern
is: every cross-reference in a docstring must be backed by a Lean
call in the body. A bridge whose body never invokes the named
target is documentation-only and rots silently as the codebase
evolves.

\item \textbf{P7 --- Defining the conclusion.} If a function is
defined as \texttt{:= <obvious target>}, a theorem that asserts
$0 \leq C/A$ for that function is trivially true. The fix is to
either push the substantive content elsewhere (a tracked-hypothesis
proposition, or a project-local infrastructure module) or commit to
a non-trivial concrete model.

\end{itemize}

The discipline is applied as a five-question prospective checklist
on each theorem statement before it is written. The reduction in
strengthening-pass cost is the primary metric, but a ruthless
post-wave review remains mandatory --- empirically, the
prospective discipline catches the more obvious failure modes
($\exists$-absorption, biconditional-tautology in conditional
violators) but misses subtler patterns: identity-function wrappers
($\theta\ h_a = h_b$), within-own-$\pm 2\sigma$-band tautologies,
pairwise-distinctness on inductive constructors, and definitional-
unfolding-as-physics. Both passes are mandatory.

The discipline is not aesthetic. Each anti-pattern corresponds to
a specific failure of the theorem-statement-as-falsifiable-claim
property: a theorem that can be discharged by \texttt{rfl} or
\texttt{decide} or \texttt{not\_lt.mpr h\_disagree} or
$0 \leq C$ for any positive $C$ does not constitute a falsifiable
physical claim. The strengthening discipline ensures that what we
prove is what we claim.

%% ─────────────────────────────────────────────────────────────────
%% §8 Sentence-level provenance + cross-paper consistency clusters
%% ─────────────────────────────────────────────────────────────────

\section{Sentence-Level Provenance and Cross-Paper Consistency Clusters}
\label{sec:provenance}

% TODO: substantive draft. Substrate to draw from:
%  - prose_state.json schema in scripts/sentence_state.py
%  - cross-paper consistency cluster mechanism in
%    scripts/cluster_detect.py + scripts/bundle_clusters.py
%  - claim_clusters.json + cluster_bundle_index.json semantics
%  - validate.py CHECK 21 (bundle_consistency) walks the registry
%    and flags cross-bundle drift
%  - The 13-bundle publication architecture per docs/PAPER_STRATEGY.md
%  - The bundle_metadata.json schema (Phase 7a sub-wave 7a.1.1) and
%    the BUNDLE_LIFT_PROCEDURE.md canonical lift workflow
%
% Section thesis: per-paper provenance tracks WHERE each numerical
% claim comes from; sentence-level provenance tracks WHAT each
% prose sentence asserts and which sentences across the program
% make the same assertion. Cross-paper consistency clusters are
% the mechanism that prevents the same fact from being stated
% differently in two papers --- a publication-readiness invariant
% that emerges when the program ships >10 papers.

\emph{Section to be drafted in subsequent session.} This section
will describe the sentence-level provenance mechanism (each prose
sentence acquires a state record tracking its source, citation
backings, and bundle destination), the cross-paper consistency
cluster registry that groups sentences asserting the same fact
across multiple papers, and the bundle architecture that
consolidates per-wave drafts into thirteen publication targets
without sacrificing per-sentence audit trails.

%% ─────────────────────────────────────────────────────────────────
%% §9 Three-layer architecture extended (Aristotle integration)
%% ─────────────────────────────────────────────────────────────────

\section{Aristotle Automated Theorem Prover Integration}
\label{sec:three_layer_extended}

% TODO: substantive draft. Substrate to draw from:
%  - paper15 §5 (Aristotle integration)
%  - The priority batching protocol
%  - Run 78dcc5f4 (25 theorems in a single session)
%  - Run 79e07d55 (G_c critical coupling)
%  - The interactive MCP loop as Stage 3a (replaces Stage 4 fallback
%    workflow described in paper15)
%
% Section thesis: Aristotle is now Stage 4 --- a fallback for
% sorries that survive the interactive Stage 3a. The interactive
% loop is roughly 1000x faster for proof construction; Aristotle
% is reserved for sorries that genuinely resist interactive proof
% with sub-lemma decomposition.

\emph{Section to be drafted in subsequent session.} This section
will describe the priority-batching protocol, the interactive Lean
language-server protocol that has displaced Aristotle as the
default proof-construction tool, and the residual role of
Aristotle as a fallback for sorries that resist interactive proof
even with sub-lemma decomposition. Empirical results from the
project's Aristotle batches will be tabulated, including the
single-run record of \aristotleproved{} cumulative theorems
across \aristotleruns{} runs.

%% ─────────────────────────────────────────────────────────────────
%% §10 Stage 13 fresh-context adversarial reviewer
%% ─────────────────────────────────────────────────────────────────

\section{The Fresh-Context Adversarial-Reviewer Pattern}
\label{sec:adversarial}

% TODO: substantive draft. Substrate to draw from:
%  - .claude/plugins/physics-qa/agents/adversarial-reviewer.md
%  - WAVE_EXECUTION_PIPELINE.md Stage 13 section
%  - The April 2026 external-adversarial-review round that
%    surfaced the 13-dimension problem space (seed event)
%  - The 8 finding classes (citation integrity, parameter
%    provenance, narrative grounding, cross-paper consistency,
%    count freshness, formal-verification-claim substantiation,
%    assumption disclosure, review-state hygiene)
%  - The supersession ledger semantics (docs/review_finding_supersessions.json)
%
% Section thesis: the adversarial reviewer runs in a fresh context
% (no memory of the author's prior justifications), works through
% 8 finding classes systematically, and produces a structured
% markdown report that flips readiness gates to blocked when
% BLOCKER findings are present. The fresh-context property is
% load-bearing --- the reviewer must not be persuadable by the
% author's prior context.

\emph{Section to be drafted in subsequent session.} This section
will describe the fresh-context property as a load-bearing design
constraint, the eight finding classes the agent works through,
the auto-pickup of findings as graph nodes, the supersession
ledger that closes findings on author fix, and the bundle-level
extension of the pattern (per-bundle Stage~13 review documents at
\texttt{papers/AutomatedReviews/<DATE>-bundle-stage13/<X>.md}).

%% ─────────────────────────────────────────────────────────────────
%% §11 Phase 6f W1-W6 first-formalization claims (substrate roster)
%% ─────────────────────────────────────────────────────────────────

\section{Algebraic-General-Relativity Substrate First-Formalization Claims}
\label{sec:phase6f_substrate}

% TODO: substantive draft. Substrate to draw from:
%  - lean/SKEFTHawking/{ADMFormalism, AlgebraicRiemann,
%    SchwarzschildKerrSchild, DeSitterAdSLambdaVacuum,
%    TetradMetricEquivalence, ...}.lean
%  - Phase 6f Wave 1-6 memory files
%  - The "first formalization in any proof assistant" claim per:
%    + algebraic Riemann (Wave 1)
%    + Schwarzschild Kerr-Schild (Wave 2)
%    + de Sitter / AdS Lambda-vacuum biconditional (Wave 3)
%    + ADM Hamiltonian/momentum-constraint biconditionals (Wave 4)
%    + tetrad metric equivalence (Wave 5)
%    + Wave 6: <topic to be filled>
%
% Section thesis: each of the six waves shipped a first-in-any-
% proof-assistant claim at the algebraic-general-relativity
% substrate level. These are not merely formalization
% achievements; they are infrastructure on which subsequent
% Phase 6 waves (BCH four laws, area theorem, Hawking-Penrose
% singularity theorem) build. The bundle paper presents this as a
% sidebar to the main methodology narrative.

\emph{Section to be drafted in subsequent session.} This section
presents a sidebar of first-formalization-in-any-proof-assistant
claims at the algebraic-general-relativity substrate level
covering algebraic Riemann curvature, Schwarzschild Kerr-Schild
form, the de Sitter / Anti-de Sitter $\Lambda$-vacuum
biconditional, the ADM Hamiltonian and momentum-constraint
biconditionals, and tetrad metric equivalence. Each is an
infrastructure result on which subsequent Phase~6 waves
(Bardeen-Carter-Hawking four laws, the area theorem, the
Hawking--Penrose singularity theorem) build.

%% ─────────────────────────────────────────────────────────────────
%% §12 Phase 6f W7 Riemannian connection sidebar
%% ─────────────────────────────────────────────────────────────────

\section{The Riemannian Connection Sidebar}
\label{sec:riemannian_sidebar}

% TODO: substantive draft. Substrate to draw from:
%  - papers/paper44_riemannian_connection/paper_draft.tex
%  - lean/SKEFTHawking/{LeviCivita, BundleRiemann,
%    BundleRiemannAux, ...}.lean
%  - The session-by-session expansion (W7 sessions 1-3 SHIPPED;
%    sessions 4-5 in flight at memory write time)
%  - The mkHom₃ + bundled CovariantDerivative.riemann construction
%  - Bundle-level first/second Bianchi via Jacobi
%  - The Mathlib upstream PR target
%
% Section thesis: the Riemannian connection sidebar documents the
% structural completion of the curvature gap above the Bonn API.
% Wave 7 Sessions 1-3 shipped the Lorentzian metric typeclass +
% Levi-Civita Christoffel uniqueness + full coordinate Riemann +
% algebraic first Bianchi + differential second Bianchi machinery.
% Wave 8 Sessions 1-3 SHIPPED add the bundle-level Riemann
% curvature consumer of Bonn IsCovariantDerivativeOn + mlieBracket
% API. The sidebar is sized for I1's methodology paper, not for
% the full 100+ pp algebraic-GR backbone paper.

\emph{Section to be drafted in subsequent session.} This section
presents the Phase~6f Wave~7 catch-up Riemannian connection
sidebar: a Lorentzian metric typeclass, the Levi-Civita
Christoffel uniqueness, the full coordinate-form Riemann tensor,
the algebraic first Bianchi identity, and the differential
second Bianchi identity. Wave~8 extends the bundle to Bonn's
\texttt{IsCovariantDerivativeOn} + \texttt{mlieBracket} API.
The sidebar is sized for the I1 methodology paper, not for the
algebraic-general-relativity backbone paper that Phase~6f
ultimately produces.

%% ─────────────────────────────────────────────────────────────────
%% §13 Phase 6g W5 structural-Prop scoping pattern
%% ─────────────────────────────────────────────────────────────────

\section{Structural-Prop Scoping Under Mathlib-Dependency Fallback}
\label{sec:structural_prop_scoping}

% TODO: substantive draft. Substrate to draw from:
%  - lean/SKEFTHawking/{HawkingPenroseSingularityCurveTheoretic,
%    AreaTheoremCurveTheoretic, RiccatiComparison, ...}.lean
%  - Phase 6g Wave 5 memory file
%  - The 1-D wave equation Fourès-Bruhat distillation case
%  - The pattern: when Mathlib lacks the PDE infrastructure
%    (Cauchy-Kowalevski, energy estimates), structure the
%    substantive content as a tracked Prop bundle that the
%    project-local proof discharges modulo the missing infra,
%    and document the missing infra as a "Mathlib upstream
%    target" rather than as a sorry
%
% Section thesis: when a substantive Lean theorem cannot be
% closed because the upstream PDE / measure-theoretic /
% functional-analytic machinery is unavailable in Mathlib, the
% structurally-correct response is not to invent the missing
% theory in-project. It is to scope the substantive content as a
% tracked Prop bundle, prove every other hypothesis in the
% bundle, and let the missing-infra Prop be the gating hypothesis
% that downstream proofs consume. This avoids in-project
% reinvention of Mathlib while keeping the substantive content
% Lean-checkable.

\emph{Section to be drafted in subsequent session.} This section
presents the Phase~6g Wave~5 structural-Prop scoping pattern: when
substantive content depends on partial-differential-equation
infrastructure that Mathlib does not yet provide
(Cauchy--Kowalevski, energy estimates, distributional Sobolev
machinery), the project's discipline is to scope the missing
infrastructure as a tracked proposition bundle, prove every other
hypothesis in the bundle, and let the missing-infrastructure
proposition be the gating hypothesis. The pattern was
exercised on the one-dimensional wave-equation Fourès-Bruhat
distillation; its absence in earlier waves caused two months of
unproductive in-project reinvention. Documenting the discipline
prevents recurrence.

%% ─────────────────────────────────────────────────────────────────
%% §14 Lessons + future work
%% ─────────────────────────────────────────────────────────────────

\section{Lessons and Future Work}
\label{sec:lessons}

We close with three lessons drawn from a year of running the
pipeline at the scale described.

\textbf{Continuous verification is cheaper than retrofitted
verification, by a margin large enough to be a workflow choice,
not a methodology preference.} Each of the four failure modes we
opened with --- a numerical claim that traces to a non-existent
code path, a ``formally verified'' claim whose theorem is a
placeholder, a citation that resolves to the wrong paper, a
parameter that has drifted from its primary source --- would
have shipped to draft state under the pre-pipeline workflow. Each
was caught, in this project, by the pipeline's continuous
checks before it could be cited downstream. The cost of running
the checks is on the order of seconds per stage; the cost of
finding and fixing a propagated error after publication is on the
order of weeks per finding, in the best case.

\textbf{Automated theorem proving is a force multiplier on
formalization work, but only when the formalization itself is
well-decomposed.} A monolithic fifty-term proof goal is
intractable to both manual proof and to Aristotle. A decomposed
nine-deep structure with each sublemma at twelve terms or fewer is
tractable to both, and Aristotle closes the sublemmas in a single
priority batch. The decomposition discipline is the pipeline's
load-bearing investment: it pays back in throughput at every
subsequent wave.

\textbf{Adversarial review is a separate role from authorship,
and the separation is load-bearing.} A reviewer in the same
context as the author has, by construction, internalized the
author's prior justifications. The reviewer's findings will, by
construction, be limited to what the author's context permits.
The fresh-context adversarial-reviewer pattern --- the reviewer
agent runs in a context window that has not seen the author's
work --- is the structural fix. The author fixes; the author
re-invokes; iteration converges to clean.

Three pipeline extensions are in active development at the time of
writing. First, a meta-process quality-improvement register
(Stage~14) that surfaces systemic patterns across waves; the seed
event is the April~2026 external-adversarial-review round that
exposed a thirteen-dimensional problem space that internal
validation could not detect by construction. Second, the
publication-bundle architecture that consolidates per-wave drafts
into thirteen publication targets while preserving sentence-level
provenance; this paper is the first bundle to be drafted under the
architecture. Third, an automated cross-bundle consistency check
that walks a cross-bundle cluster registry and flags member-
sentence drift across bundle boundaries, preventing the same fact
from being stated differently in two papers.

The pipeline is open-source, domain-agnostic, and immediately
applicable in any research domain where the connection between
computation and published claims must be trustworthy. We
specifically encourage adoption in any setting where the cost of a
correctness failure post-publication exceeds the cost of running
continuous verification --- a condition that holds in
pharmaceutical software (FDA SaMD algorithmic vigilance),
verified-risk financial modeling, climate modeling, and any other
research domain where the audit trail from code to claim is itself
the deliverable.

%% ─────────────────────────────────────────────────────────────────

\begin{thebibliography}{99}
\bibitem{physlean} J.~Tooby-Smith, ``PhysLean: Lean~4 for physics,''
\url{https://github.com/HEPLean/PhysLean} (2024).
\bibitem{compcert} X.~Leroy, ``Formal verification of a realistic
compiler,'' \emph{Comm.\ ACM}\ \textbf{52}(7), 107--115 (2009).
\bibitem{sel4} G.~Klein et al., ``seL4: Formal verification of an OS
kernel,'' \emph{Proc.\ ACM SOSP} (2009).
\bibitem{leancipilot} P.~Song et al., ``Towards large language models
as copilots for theorem proving in Lean,'' arXiv:2404.12534 (2024).
\bibitem{alphaproof} Google DeepMind, ``AI achieves silver-medal
standard solving International Mathematical Olympiad problems,''
(2024).
\bibitem{paper15methodology} J.~Roehm, ``A Twelve-Stage Verification
Pipeline for Computational Physics Research'' (in prep, project
internal draft).
\bibitem{paper44riemannian} J.~Roehm, ``Riemannian Connection on the
Algebraic-General-Relativity Substrate'' (in prep, project internal
draft).
\end{thebibliography}

\end{document}
