\documentclass[aps,prd,twocolumn,showpacs,superscriptaddress,nofootinbib]{revtex4-2}

\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{hyperref}

\begin{document}

\title{Spectral Predictions for Analog Hawking Radiation:\\
Platform-Specific Tables from SK-EFT}

\author{SK-EFT Hawking Research Program}

\date{\today}

\begin{abstract}
We present platform-specific spectral predictions for analog Hawking
radiation in three BEC experiments: Steinhauer $^{87}$Rb (Technion),
Heidelberg $^{39}$K, and Trento $^{23}$Na spin-sonic. The predictions
derive from the Schwinger-Keldysh dissipative effective field theory
framework (Papers~1--4) and include: (i)~complete spectral occupation
$n(\omega)$ at representative frequencies, (ii)~correction decomposition
into dispersive and dissipative contributions, (iii)~noise floor
thresholds from the fluctuation-dissipation relation, (iv)~detector
sensitivity requirements for three measurement goals, and
(v)~a kappa-scaling test prediction for Heidelberg~$^{39}$K.
This document is designed for experimentalists planning measurements.
\end{abstract}

\maketitle

\section{Platform Parameters}

\begin{table}[h]
\caption{BEC platform parameters in natural units ($c_s = 1$, $\kappa = 1$).}
\label{tab:params}
\begin{ruledtabular}
\begin{tabular}{lccc}
Parameter & Steinhauer & Heidelberg & Trento \\
\hline
Atom & $^{87}$Rb & $^{39}$K & $^{23}$Na \\
$D$ & 0.030 & 0.020 & 0.014 \\
$\gamma_\text{dim}$ & $3\times10^{-3}$ & $2\times10^{-3}$ & $1.4\times10^{-5}$ \\
$T_H$ & $\kappa/2\pi$ & $\kappa/2\pi$ & $\kappa/2\pi$ \\
$\omega_\text{max}/T_H$ & 65.1 & 85.3 & 108.2 \\
\end{tabular}
\end{ruledtabular}
\end{table}

\section{Spectral Predictions}

Table~\ref{tab:spectral} lists the predicted occupation number $n(\omega)$
at five representative frequencies for each platform, along with the
decomposition into dispersive ($\delta_\text{disp}$) and dissipative
($\delta_\text{diss}$) corrections.

The total occupation is:
\begin{equation}
n(\omega) = n_\text{Hawking}(\omega) + n_\text{noise}(\omega)
\end{equation}
where $n_\text{Hawking} = |\beta|^2 / (1 - \delta_k)$ includes the
modified unitarity correction and $n_\text{noise} = \delta_k/2$ is the
FDR-mandated noise floor.

\begin{table*}[t]
\caption{Spectral predictions at representative frequencies. Values computed
from the exact WKB connection formula (Paper~4).}
\label{tab:spectral}
\begin{ruledtabular}
\begin{tabular}{l c c c c c c c}
Platform & $\omega/T_H$ & $n_\text{total}$ & $n_\text{Planck}$ & $\delta_\text{disp}$ & $\delta_\text{diss}$ & $\delta_k$ & Dominant \\
\hline
\multicolumn{8}{c}{\textit{Steinhauer $^{87}$Rb}} \\
 & 0.5 & 1.903 & 1.902 & $-4.7\times10^{-4}$ & $1.4\times10^{-5}$ & $2.7\times10^{-5}$ & dispersive \\
 & 1.0 & 0.525 & 0.524 & $-4.7\times10^{-4}$ & $8.7\times10^{-5}$ & $1.7\times10^{-4}$ & dispersive \\
 & 2.0 & 0.151 & 0.150 & $-4.7\times10^{-4}$ & $3.2\times10^{-4}$ & $6.3\times10^{-4}$ & noise floor \\
 & 3.0 & 0.053 & 0.052 & $-4.7\times10^{-4}$ & $6.9\times10^{-4}$ & $1.4\times10^{-3}$ & noise floor \\
 & 5.0 & 0.009 & 0.007 & $-4.7\times10^{-4}$ & $1.9\times10^{-3}$ & $3.7\times10^{-3}$ & noise floor \\
\hline
\multicolumn{8}{c}{\textit{Heidelberg $^{39}$K}} \\
 & 0.5 & 1.903 & 1.902 & $-2.1\times10^{-4}$ & $9.0\times10^{-6}$ & $1.8\times10^{-5}$ & dispersive \\
 & 1.0 & 0.524 & 0.524 & $-2.1\times10^{-4}$ & $5.8\times10^{-5}$ & $1.2\times10^{-4}$ & dispersive \\
 & 2.0 & 0.151 & 0.150 & $-2.1\times10^{-4}$ & $2.1\times10^{-4}$ & $4.2\times10^{-4}$ & noise floor \\
 & 3.0 & 0.053 & 0.052 & $-2.1\times10^{-4}$ & $4.6\times10^{-4}$ & $9.1\times10^{-4}$ & noise floor \\
 & 5.0 & 0.008 & 0.007 & $-2.1\times10^{-4}$ & $1.2\times10^{-3}$ & $2.5\times10^{-3}$ & noise floor \\
\hline
\multicolumn{8}{c}{\textit{Trento $^{23}$Na}} \\
 & 0.5 & 1.903 & 1.902 & $-1.0\times10^{-4}$ & $6.3\times10^{-8}$ & $1.3\times10^{-7}$ & dispersive \\
 & 1.0 & 0.524 & 0.524 & $-1.0\times10^{-4}$ & $4.0\times10^{-7}$ & $8.1\times10^{-7}$ & dispersive \\
 & 2.0 & 0.150 & 0.150 & $-1.0\times10^{-4}$ & $1.5\times10^{-6}$ & $2.9\times10^{-6}$ & dispersive \\
 & 3.0 & 0.052 & 0.052 & $-1.0\times10^{-4}$ & $3.2\times10^{-6}$ & $6.4\times10^{-6}$ & dispersive \\
 & 5.0 & 0.007 & 0.007 & $-1.0\times10^{-4}$ & $8.6\times10^{-6}$ & $1.7\times10^{-5}$ & noise floor \\
\end{tabular}
\end{ruledtabular}
\end{table*}

The key physics features visible in Table~\ref{tab:spectral}:
\begin{itemize}
\item At low frequencies ($\omega \lesssim T_H$), the dispersive correction
dominates. It is negative (subluminal dispersion lowers $T_\text{eff}$).
\item At high frequencies ($\omega \gtrsim 2\,T_H$), the noise floor
dominates for Steinhauer and Heidelberg due to their larger damping rates.
\item Trento has negligible dissipation ($\gamma_\text{dim} = 1.4\times10^{-5}$),
so dispersive corrections dominate across almost the entire spectrum.
\item The crossover from dispersive to noise-floor dominance occurs at
$\omega/T_H \approx 2$ for Steinhauer and Heidelberg.
\end{itemize}

\section{Detector Requirements}

Three measurement goals are considered:
\begin{enumerate}
\item Detect the first-order dissipative correction $\delta_\text{diss}$
at 3$\sigma$ significance.
\item Detect the FDR noise floor $n_\text{noise}$ at 3$\sigma$.
\item Distinguish exact WKB from perturbative EFT at the 1\% level.
\end{enumerate}

\section{Kappa-Scaling Test}

The Heidelberg $^{39}$K platform can tune the scattering length via
Feshbach resonance, enabling a direct test of how the Hawking spectrum
scales with surface gravity $\kappa$.

The SK-EFT prediction provides a clean discriminant:
\begin{itemize}
\item $\delta_\text{disp} \propto D^2 \propto \kappa^2$ (quadratic scaling)
\item $\delta_\text{diss} \approx \text{const}$ (independent of $\kappa$
at leading order)
\end{itemize}

Measuring the $\kappa$-dependence of the spectral deviation separates
dispersive from dissipative corrections without needing absolute
calibration of either.

\section{Summary: Which Platform for Which Measurement}

\begin{table}[h]
\caption{Platform rankings by measurement goal.}
\label{tab:rankings}
\begin{ruledtabular}
\begin{tabular}{lc}
Measurement & Best Platform \\
\hline
Initial EFT detection & Steinhauer \\
FDR noise floor & Steinhauer \\
UV spectral range & Trento \\
Kappa scaling & Heidelberg \\
Fewest shots needed & Steinhauer \\
\end{tabular}
\end{ruledtabular}
\end{table}

\end{document}
