\documentclass[aps,prd,reprint,nofootinbib,superscriptaddress]{revtex4-2}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{braket}

%% Bundle: D1 (tier 1, target PRD long article)
%% Source manifest: papers/D1/source_manifest.md
%% Change log: papers/D1/change_log.md
%% Phase 7b sub-wave 7b.6 — D1 lift (skeleton + first-pass substantive draft)
%% Sources: paper1+paper2+paper4+paper12+paper16_graphene_sk_eft
%% E1 PRL (polariton) and E2 PRL (graphene) extracted Tier-4 letters
%% already shipped GREEN (Phase 7b sub-wave 7b.4, commit ea1d944)

\begin{document}

\title{Formally Verified Analog Hawking Radiation Across Three Platforms}

\author{John Roehm}
\affiliation{Independent Researcher}
\email{jgroehm@gmail.com}

\date{\today}

\begin{abstract}
We present a unified, formally verified treatment of analog Hawking
radiation across three condensed-matter platforms: atomic
Bose-Einstein condensates, exciton-polariton microcavities, and
bilayer-graphene Dirac fluids. The Schwinger-Keldysh effective field
theory (SK-EFT) framework is presented to all orders, with the
transport-coefficient counting formula
$\mathrm{count}(N) = \lfloor (N{+}1)/2\rfloor + 1$ proved by
Aristotle (run \texttt{d61290fd}). Positivity at second order forces
$\gamma_{2,1} + \gamma_{2,2} = 0$, reducing four free parameters to
three; KMS optimality is closed by Aristotle
(\texttt{SKDoubling.firstOrder\_KMS\_optimal}). The exact WKB connection
formula reveals three non-perturbative phenomena absent at any
finite order in perturbation theory: modified unitarity
$|\alpha|^2 - |\beta|^2 = 1 - \delta_k$ with
$\delta_k = 2\Gamma_H/\kappa$, an FDR-mandated noise floor
$n_\text{noise} = \delta_k / 2$ that is mandatory by KMS even with
zero Hawking radiation, and a spectral floor at
$\omega_\times \approx 6\, T_H$ where the high-$\omega$ spectrum
becomes power-law rather than thermal. The three platforms span ten
orders of magnitude in Hawking temperature
($T_H \sim \SI{0.35}{\nano\kelvin}$ for $^{87}$Rb BEC,
$T_H \sim \SI{85}{\milli\kelvin}$ for the LKB polariton smooth-horizon
baseline, $T_H \approx \SI{2.4}{\kelvin}$ for the Dean bilayer
graphene nozzle), yet share \emph{$92\%$ of the Lean theorem
infrastructure} (109 of 119 1+1D theorems apply unchanged to
graphene's 2+1D Dirac-fluid setup with $c_s \to v_F/\sqrt{2}$). The
hierarchy of corrections is platform-dependent: at Heidelberg-class
BEC the dissipative correction $\delta_\text{diss}$ exceeds the
dispersive $\delta_\text{disp}$ by a factor $\sim 22$; at the LKB
polariton smooth-horizon configuration $\delta_\text{disp}$ is
$\approx -19\%$ and dominant; at the Dean graphene nozzle
$\delta_\text{disp} \approx -2.8\%$ exceeds
$\delta_\text{diss} \sim 10^{-13}$ by eleven orders of magnitude.
Each platform yields a distinct leading detection mode: BEC delivers
the FDR noise floor and (at Heidelberg) the spectral-floor
crossover; polariton delivers stimulated-Hawking gain
$G(\omega) > 0.5$ at the universal frequency
$\omega/\kappa = \ln(3)/(2\pi) \approx 0.175$ accessible at
$N_\text{probe} \sim 100$ probe photons per mode (factor
$10^3$--$10^6$ improvement over spontaneous-correlation detection);
graphene delivers a closed-form noise spectrum
$\Delta S_I(\omega) = 2\hbar\omega \sigma_Q \Gamma(\omega) n_H(\omega)$
derived independently from Keldysh fluctuation-dissipation and
Landauer-Büttiker scattering, validated cross-platform by the
giant Wiedemann-Franz violation $L/L_0 > 200$ measured at the same
device. End-to-end, $5229$ Lean theorems are machine-verified
(\texttt{lake build} clean, zero \texttt{sorry}, $322$ closed by the
Aristotle automated prover). The single substantive axiom in the
chain (\texttt{FirstOrderKMS}) was reformulated after Aristotle's
counterexample $c = (0,0,0,0,0,0,0,1,0)$ exposed that the original
axiom constrained only four of nine field components --- a
methodological showcase of the value of formal verification for
EFT axiom design. Companion experimental letters E1 (Paris-LKB
polariton) and E2 (Dean-Kim-Lucas graphene) carry the
experimental-team-targeted implementations.
\end{abstract}

\maketitle

%% ══════════════════════════════════════════════════════════════════
\section{Introduction}
\label{sec:intro}

Analog Hawking radiation has been an experimental program for half
a century~\cite{Unruh1981,Visser1998,Hawking1974}. Three independent
platforms have now reached the regime in which a fluid horizon and
its corresponding Hawking spectrum can be plausibly resolved: atomic
$^{87}$Rb Bose-Einstein
condensates~\cite{Steinhauer2016,deNova2019}, exciton-polariton
microcavities~\cite{Falque2025,Giacobino2025}, and bilayer-graphene
Dirac fluids in supersonic nozzle
geometries~\cite{Geurs2025,Majumdar2025}. The three platforms span a
ten-order-of-magnitude window in Hawking temperature
($\SI{0.35}{\nano\kelvin}$ to $\SI{2.4}{\kelvin}$) and use
microscopic constituents that differ as profoundly as cold atoms
versus driven exciton-photons versus relativistic fermions. The
question this paper answers is: \emph{does the analog Hawking
prediction depend on these microscopic differences, or is it a
universal consequence of horizon kinematics in any fluid?}

We make the universality claim quantitative through formal
verification. We present an SK-EFT
framework~\cite{CGL2017a,CGL2017b,Crossley2017,Jana2020} for the
three platforms and machine-check its predictions in Lean~4. The
universality claim becomes the statement that 109 of 119 BEC 1+1D
Lean theorems apply unchanged to the graphene 2+1D Dirac-fluid
setup with $c_s \to v_F/\sqrt{2}$ --- a $92\%$ reuse fraction. The
$8\%$ that does not reuse ($\sim 10$ theorems) handles the metric
block-diagonalisation specific to the 2+1D quasi-1D flow geometry.
This is, to our knowledge, the strongest quantitative test of
analog-Hawking universality across distinct microscopic substrates.

The headline contributions of this paper are not any one platform's
prediction in isolation: per-wave drafts of (BEC SK-EFT first
order)~\cite{Roehm2026Wave1}, (BEC SK-EFT second
order)~\cite{Roehm2026Wave2}, (exact WKB
connection)~\cite{Roehm2026Wave4}, (polariton
predictions)~\cite{Roehm2026Wave12}, and (graphene Dirac-fluid
predictions)~\cite{Roehm2026Wave16} exist independently. What is new
here is the unified architecture: same SK-EFT machine, three
platforms, three distinct leading observables. The synthesis-level
claim is the cross-platform universality itself, made quantitative
through Lean-theorem reuse and made falsifiable through the three
non-perturbative predictions of the exact-WKB connection formula.

This paper is organised as follows. Section~\ref{sec:bec-first}
presents the first-order SK-EFT in the BEC reference setting:
Son's superfluid action, SK contour doubling, three axioms (Galilean
invariance, gradient expansion, KMS positivity), two transport
coefficients $(\gamma_1, \gamma_2)$, and the foundational closed
form $\delta_\text{diss} = (\gamma_1 + \gamma_2)\kappa/c_s^2$.
Section~\ref{sec:bec-second} extends to all orders via the counting
formula $\mathrm{count}(N) = \lfloor (N{+}1)/2 \rfloor + 1$
(Aristotle), shows that positivity-forced relations reduce
$4 \to 3$ free parameters at second order, derives the CGL
fluctuation-dissipation relation
$K_N(\omega,k) = -\mathrm{i}\,[K_R(\omega) - K_R(-\omega)]/(\beta\omega)$
from algebraic KMS, proves KMS optimality, and identifies the
parity-asymmetry signature that separates dispersive and
dissipative corrections experimentally.
Section~\ref{sec:exact-wkb} is the workhorse: the exact-WKB
connection formula and the three non-perturbative effects.
Section~\ref{sec:hierarchy} closes the BEC chain with the
hierarchy of corrections by platform.
Section~\ref{sec:polariton} treats the polariton platform,
including the reservoir-corrected sound speed, the
stimulated-Hawking detection, and the universal threshold
$\omega/\kappa = \ln(3)/(2\pi) \approx 0.175$ at $G(\omega) = 0.5$.
Section~\ref{sec:graphene} treats the graphene Dirac-fluid platform:
the 3$\times$3 metric block-diagonalisation, the $92\%$ Lean theorem
reuse claim, the closed-form noise spectrum from independent FDT
and Landauer-Büttiker derivations, and the Wiedemann-Franz
cross-check linking the SK-EFT noise prediction to a published
transport measurement. Section~\ref{sec:synthesis} synthesises the
cross-platform picture, recapitulates the Aristotle-counterexample
sidebar as a methodology showcase, and sets up the open questions
that the Phase-6+ programme will address.

\textbf{Companion papers.} The experimental letters
E1~\cite{Roehm2026E1} (Paris-LKB polariton) and
E2~\cite{Roehm2026E2} (Dean-Kim-Lucas graphene) ship as Tier-4
PRL/PRR letters paired with this deep paper. They contain the
experimental-team-targeted predictions; this paper supplies the
theory and Lean backbone. The methodology paper
I1~\cite{Roehm2026I1} elaborates the verification pipeline.

%% ══════════════════════════════════════════════════════════════════
\section{First-order SK-EFT: the BEC reference}
\label{sec:bec-first}

\subsection{Son's superfluid action and SK doubling}

The starting point is Son's superfluid effective
action~\cite{Son2002}, which captures the Goldstone mode $\phi$ of
broken U(1) particle-number symmetry on the acoustic metric
$g_{\mu\nu}^{\rm ac}$ produced by a transonic background flow.
Promoting Son's action to the closed-time-path Schwinger-Keldysh
contour~\cite{CGL2017a,CGL2017b} doubles each field
$\phi \to (\phi_1, \phi_2)$ and adds a bilinear noise vertex.
The SK action takes the Crossley-Glorioso-Liu (CGL)
form~\cite{Crossley2017},
\begin{equation}
S_\text{SK}[\phi_+, \phi_-]
\;=\; \int dt\,d^d x \, \mathcal{L}_\text{SK}[\phi_+, \phi_-]
\end{equation}
with $\phi_\pm = (\phi_1 \pm \phi_2)/\sqrt{2}$ Keldysh-rotated
fields. The advantage of the CGL form is that the quadratic part
splits cleanly into retarded ($K_R$) and noise ($K_N$) sectors,
with the dynamical KMS condition imposing
$K_N = -\mathrm{i}[K_R - K_R^*]/(\beta\omega)$ in the linear
regime~\cite{Crossley2017}.

\subsection{Three SK-EFT axioms and two transport coefficients}

For analog gravity in non-relativistic superfluids, the SK-EFT is
constrained by three axioms encoded in our Lean library: Galilean
invariance of the underlying superfluid; gradient expansion in
powers of $\partial/(c_s\, k_\text{UV})$; KMS positivity (the
noise coefficients $\gamma_i$ are non-negative). The KMS structure is
encoded as the \texttt{FirstOrderKMS} record in
\texttt{SecondOrderSK.lean} with fluctuation-dissipation fields
\texttt{fdr\_i1}, \texttt{fdr\_i2}, \texttt{i3\_zero}; the original
axiom was reformulated as a Prop bundle after the Aristotle
counterexample (see~\S\ref{sec:synthesis}).

At first order in the gradient expansion there are exactly two
positive transport coefficients $\gamma_1, \gamma_2$ (bulk and
shear sound damping). The leading dissipative correction to the
Hawking spectrum is
\begin{equation}
\delta_\text{diss}(\omega) \;=\; \frac{(\gamma_1 + \gamma_2)\,\kappa}{c_s^2}
\,\equiv\, \frac{\Gamma_H}{\kappa},
\label{eq:delta-diss}
\end{equation}
where the Hawking damping rate $\Gamma_H = (\gamma_1+\gamma_2)\kappa^2/c_s^2$
is the IR-dominant scale. This is captured at the Lean level in
\texttt{SecondOrderSK.deltaDissFromTransport\_eq}.

\subsection{BEC platform numerics}

The hierarchy is platform-dependent and culminates in
Heidelberg-class spin-sonic configurations where
$\delta_\text{diss} \sim 10^{-1}$ is the dominant correction
(see~\S\ref{sec:hierarchy} for the full table). The Stage-13 anchor
$\delta_\text{diss} = (1 + \xi^2)^{-1}$ at the canonical BEC anchor
parameters (\texttt{EXPERIMENTS['Steinhauer']} as registered in
\texttt{src/core/constants.py}; the corresponding Heidelberg
parameters are similarly registered) is encoded in the formula
chain rooted at \eqref{eq:delta-diss}; the $\xi^2$ factor absorbs
the Bogoliubov healing-length contribution to the bulk modulus.

\subsection{Acoustic-metric infrastructure}

Four Lean theorems comprise the acoustic-metric backbone:
\texttt{AcousticMetric.acousticMetric\_det},
\texttt{AcousticMetric.acoustic\_metric\_lorentzian},
\texttt{AcousticMetric.acousticMetric\_inv\_correct}, and
\texttt{hawking\_temp\_from\_surface\_gravity}. These are reused
unchanged in the 2+1D graphene context (with $c_s \to v_F/\sqrt{2}$);
they form the backbone of the cross-platform reuse claim of
\S\ref{sec:graphene}.

%% ══════════════════════════════════════════════════════════════════
\section{Second-order SK-EFT: counting, KMS optimality, and parity}
\label{sec:bec-second}

\subsection{Transport coefficient counting}

A central technical result is that the number of independent
transport coefficients at order $N$ in the gradient expansion is
\begin{equation}
\mathrm{count}(N) \;=\; \left\lfloor \frac{N+1}{2} \right\rfloor + 1,
\label{eq:count-formula}
\end{equation}
giving $\mathrm{count}(1) = 2$, $\mathrm{count}(2) = 2$ (without
parity asymmetry) or $\mathrm{count}(3) = 3$, etc. The proof, by
explicit bijection to a parity-graded monomial set, is closed by
the Aristotle automated prover (run \texttt{d61290fd}); the
small-$N$ cases $N = 1, 2, 3$ are independently closed by
\texttt{native\_decide} in
\texttt{SecondOrderSK.firstOrder\_count},
\texttt{secondOrder\_count},
\texttt{secondOrder\_count\_with\_parity},
\texttt{thirdOrder\_count}.

\subsection{Positivity-forced relations}

At second order, KMS positivity forces a single linear relation
between the two new coefficients:
\[
\gamma_{2,1} + \gamma_{2,2} \;=\; 0.
\]
This reduces the apparent four free parameters
$(\gamma_1, \gamma_2, \gamma_{2,1}, \gamma_{2,2})$ to three.
The reduction is captured in
\texttt{SecondOrderSK.fullSecondOrder\_uniqueness}. Both new
second-order coefficients require broken spatial parity: under
$x \to -x$ symmetry they vanish identically. This is the cleanest
experimental signature separating dispersive corrections (which
survive parity) from dissipative ones (which do not) at second
order.

\subsection{KMS optimality}

The KMS positivity axiom is optimal in the sense that no further
relations among the transport coefficients are imposed. Formally,
positivity is equivalent to $i_1 \geq 0 \wedge i_2 \geq 0$ where
$(i_1, i_2)$ are the eigenvalues of the spectral matrix; no hidden
inequalities exist. This is closed by Aristotle in
\texttt{SKDoubling.firstOrder\_KMS\_optimal}. The companion stress
test \texttt{SKDoubling.firstOrder\_altSign\_uniqueness\_test} verifies
the optimality numerically by exhibiting parameter-space points that
fail when the axiom is relaxed.

\subsection{The CGL fluctuation-dissipation relation}

Combining algebraic KMS with the second-order action yields the
master FDR
\begin{equation}
K_N(\omega, \mathbf{k})
\;=\; -\mathrm{i} \cdot \frac{K_R(\omega, \mathbf{k}) - K_R(-\omega, \mathbf{k})}{\beta\,\omega},
\label{eq:cgl-fdr}
\end{equation}
which connects the algebraic SK axioms to the dynamical-KMS
microscopic principle of Crossley-Glorioso-Liu. The chain of
implications --- algebraic KMS $\Rightarrow$ Einstein relation
$\Rightarrow$ CGL FDR $\Rightarrow$ second-order KMS --- is closed
in \texttt{CGLTransform.einstein\_relation},
\texttt{cgl\_fdr\_general}, and
\texttt{cgl\_implies\_secondOrderKMS}. The CGL FDR is what makes
the KMS-mandated noise floor of \S\ref{sec:exact-wkb} a
\emph{theorem}, not an assumption.

\subsection{Acoustic-shell vanishing of $\delta^{(2)}$}

A key result of the second-order analysis: the second-order
correction $\delta^{(2)}(\omega)$ vanishes identically on the
acoustic shell ($\omega^2 = c_s^2 k^2$). It becomes nonzero only
through Bogoliubov off-shell dispersion at order
$\xi^2 k^2 / (c_s\, k_\text{UV})^2$. For all current BEC platforms,
$\delta^{(2)}$ is at most $10^{-8}$--$10^{-12}$ --- below detection
thresholds. This is captured in the second-order formal-claim
infrastructure of \texttt{SecondOrderSK}.

%% ══════════════════════════════════════════════════════════════════
\section{Exact WKB connection: three non-perturbative effects}
\label{sec:exact-wkb}

This section is the workhorse. The exact-WKB connection
formula~\cite{Berry1989,Heading1962,CorleyJacobson1996,
CoutantParentani2012,RobertsonParentani2015,BelgiornoCacciatori2024}
provides three predictions absent at any finite order in
perturbation theory.

\subsection{Complex turning-point shift}

The dissipative correction shifts the WKB turning point off the
real axis by
\begin{equation}
\delta x_\text{imag} \;=\; \frac{c_s\, \Gamma_H}{2\,\kappa^2}.
\label{eq:turning-point-shift}
\end{equation}
This is exact to all orders in the SK-EFT; the EFT order enters
only through $\Gamma_H(\omega)$. The Lean
\texttt{WKBConnection.turning\_point\_shift\_pos\_iff}
formalises the equivalence of this shift to the dissipative
correction being non-zero.

\subsection{Modified Bogoliubov coefficients}

Standing-wave matching across the complex turning point gives
\begin{equation}
\left| \frac{\beta(\omega)}{\alpha(\omega)} \right|^2
\;=\; \exp\!\left( -\frac{2\pi\omega}{\kappa_\text{eff}(\omega)} \right),
\quad
\kappa_\text{eff}(\omega) = \frac{\kappa}{1 + \delta_\text{disp} + \delta_\text{diss}(\omega)}.
\end{equation}
The recovery of standard unitarity in the $\Gamma_H \to 0$ limit is
captured by \texttt{WKBConnection.standard\_unitarity\_recovered}.

\subsection{Non-perturbative effect 1 --- modified unitarity}

Generalised across the dissipative SK contour, the standard
relation $|\alpha|^2 - |\beta|^2 = 1$ is replaced by
\begin{equation}
|\alpha(\omega)|^2 - |\beta(\omega)|^2 \;=\; 1 - \delta_k,
\qquad
\delta_k \;=\; \frac{2\,\Gamma_H}{\kappa},
\label{eq:modified-unitarity}
\end{equation}
where $\delta_k$ is the decoherence parameter. This is encoded in
\texttt{WKBConnection.unitarity\_deficit\_eq\_decoherence};
non-negativity in
\texttt{WKBConnection.decoherence\_nonneg}; the
$\Gamma_H \to 0$ recovery in
\texttt{WKBConnection.decoherence\_zero\_iff}. Physically: the
final state of the analog Hawking process is not a pure quantum
state; environmental coupling forces a finite mixed-state weight.

\subsection{Non-perturbative effect 2 --- FDR-mandated noise floor}

Even with $|\beta|^2 = 0$ (no Hawking radiation), the FDR
\eqref{eq:cgl-fdr} forces a residual
\begin{equation}
n_\text{noise} \;=\; \frac{\delta_k}{2} \;=\; \frac{\Gamma_H}{\kappa},
\label{eq:noise-floor}
\end{equation}
encoded in \texttt{WKBConnection.noise\_floor\_nonneg} and
\texttt{noise\_floor\_eq\_delta\_diss}. This is mandatory by KMS:
the FDR cannot be turned off without violating algebraic axioms of
the SK-EFT. The noise floor is observable as a flat baseline to
which the Hawking spectrum sits as a Planckian increment.

\subsection{Non-perturbative effect 3 --- spectral floor}

Combining \eqref{eq:modified-unitarity} and \eqref{eq:noise-floor}
gives the total occupation number
\begin{equation}
n(\omega) \;=\; |\beta(\omega)|^2 \;+\; \frac{\delta_k}{2}.
\end{equation}
Encoded in \texttt{WKBConnection.occupation\_reduces\_to\_beta\_sq}.
The Hawking term decays as $\exp(-\omega/T_H)$ at high frequency,
while the noise floor is constant. The crossover frequency, where
\texttt{WKBConnection.noise\_dominates\_when\_beta\_small}, is at
\[
\omega_\times \;\approx\; T_H \cdot \ln(2 / \delta_k).
\]
For Heidelberg-class BEC ($\delta_\text{diss} \approx 0.26$;
Tab.~\ref{tab:hierarchy}), $\omega_\times \approx T_H \ln(2/0.26)
\approx 2.0\, T_H$ --- well below the dispersive UV cutoff. The
high-frequency spectrum becomes power-law ($\omega^{-1}$ correction)
rather than exponentially suppressed. This is, to our knowledge, the
strongest-yet \emph{universal} prediction of dissipative analog
Hawking radiation: a regime where the spectrum is thermal at
$\omega \lesssim T_H$, transitions to a Planckian floor at
$\omega \sim T_H$, and bottoms out at the $\omega^{-1}$ spectral
floor at $\omega \gtrsim 2\, T_H$. (For Steinhauer-class
$\delta_\text{diss} \approx 10^{-5}$ the crossover sits at
$\omega_\times \approx T_H \ln(2\!\times\!10^{5}) \approx 12\, T_H$,
well above the dispersive UV cutoff.)

\subsection{Heidelberg as the unique BEC target}

Among the three BEC platforms, only Heidelberg sits in the regime
where the spectral floor is below the dispersive UV cutoff.
Steinhauer and Trento have $\omega_\times$ above the dispersive
cutoff; the spectral floor exists in principle but cannot be
probed without extending the platform.

%% ══════════════════════════════════════════════════════════════════
\section{Hierarchy of corrections by platform}
\label{sec:hierarchy}

The hierarchy of dispersive vs.~dissipative corrections varies
strongly by platform. Table~\ref{tab:hierarchy} summarises.

\begin{table*}[!t]
\centering
\caption{Hierarchy of corrections at three benchmark platforms.
$\delta_\text{disp}$ is the leading dispersive correction
($-\pi D^2/6$ for sub-luminal dispersion);
$\delta_\text{diss}$ is from~\eqref{eq:delta-diss};
$\delta^{(2)}$ is the on-shell second-order correction. Numerical
values are computed via \texttt{src/core/formulas.py} on
\texttt{EXPERIMENTS} parameters.}
\label{tab:hierarchy}
\begin{tabular}{lcccc}\toprule
Platform & $T_H$ & $\delta_\text{disp}$ & $\delta_\text{diss}$ & dominance \\\midrule
$^{87}$Rb BEC (Steinhauer) & $\sim 0.35$\,nK & $-2.7\%$ & $10^{-5}$ & disp \\
$^{87}$Rb BEC (Heidelberg) & $\sim 1$\,nK & $-1.2\%$ & $\sim 26\%$ & \textbf{diss} \\
spin-sonic BEC (Trento)    & $\sim 1$\,nK & $-0.5\%$ & $10^{-1}$ & comparable \\
LKB polariton (smooth)     & $\sim 85$\,mK & $-19\%$ & $14\%$ & disp/comp \\
LKB polariton (steep)      & $\sim 134$\,mK & $-46\%$ & $\sim 10\%$ & \textbf{disp} (LO only) \\
graphene (Dean)            & $\sim 2.4$\,K & $-2.8\%$ & $\sim 10^{-13}$ & \textbf{disp} \\
\bottomrule
\end{tabular}
\end{table*}

The hierarchy inversion at Heidelberg ($\delta_\text{diss}$ exceeds
$\delta_\text{disp}$ by factor $\sim 22$) is the single most
accessible BEC test of the SK-EFT framework. The platform-specific
scaling $\Gamma_\text{Beliaev} \propto \kappa^2$ versus
$\delta_\text{disp} \propto \kappa^2$ implies that the ratio
$\delta_\text{diss}/\delta_\text{disp}$ is independent of $\kappa$
within a fixed platform but varies by orders of magnitude between
platforms. The most diagnostic experimental signature is therefore
not the absolute value of $\delta_\text{diss}$ but its
$\kappa$-scaling relative to $\delta_\text{disp}$, accessible at
Heidelberg through Feshbach-tuned scattering-length sweeps.

%% ══════════════════════════════════════════════════════════════════
\section{Polariton platform: stimulated Hawking and reservoir
correction}
\label{sec:polariton}

The polariton platform realises analog Hawking at $T_H \sim 85$\,mK,
roughly $10^8$ above atomic BEC. The driven-dissipative nature of
the platform alters the noise structure (effective polariton
temperature $T_\text{noise} \sim 0.81$\,K dominates $T_H$) but
introduces a new detection mode: stimulated Hawking emission.

\subsection{Reservoir-corrected sound speed}

Falque \emph{et al.}~\cite{Falque2025} measure
$c_s \approx 0.40$\,$\mu$m/ps directly via shock-front velocimetry
across multiple condensate densities, factor $\sim 2.5$ below
reservoir-free theoretical estimates. The discrepancy is explained
by the excitonic reservoir absorbing $\sim 50$--$75\%$ of the
condensate blueshift without contributing dynamic sound speed.
Three independent measurements~\cite{Estrecho2021,Stepanov2019,Claude2023}
confirm the reservoir correction. We adopt the Falque measurement
directly throughout.

\subsection{Smooth- and steep-horizon configurations}

The LKB platform supports a smooth-horizon configuration with
$\kappa = \SI{7e10}{\per\second}$, $D \approx 0.60$, and a
steep-horizon reach with $\kappa = \SI{1.1e11}{\per\second}$,
$D \approx 0.93$. Hawking temperatures are
$T_H \approx 85$\,mK and $T_H \approx 134$\,mK respectively. The
smooth-horizon configuration sits at the borderline between the
perturbative and borderline regimes
($\Gamma_\text{pol}/\kappa \approx 0.143$); the steep-horizon
configuration is leading-order only ($D > 0.5$).

\subsection{Stimulated-Hawking detection: the universal threshold}

A coherent probe at frequency $\omega$ incident on the analog
horizon stimulates emission with gain
\begin{equation}
G(\omega) \;=\; \frac{\Gamma(\omega)}{e^{2\pi\omega/\kappa_\text{eff}} - 1}.
\end{equation}
At single-shot $5\sigma$ detection threshold, $G > 0.5$
(equivalently $N_\text{probe} = 25/G^2 = 100$ probe photons per
mode). The threshold is crossed at
\begin{equation}
\boxed{\;\omega/\kappa \;<\; \frac{\ln 3}{2\pi} \;\approx\; 0.175\;}
\label{eq:stim-threshold}
\end{equation}
which is universal and independent of $\kappa$. The Lean
infrastructure in \texttt{PolaritonTier1} provides the foundational
theorems \texttt{attenuation\_ge\_one} (spatial attenuation
$\geq 1$ everywhere), \texttt{attenuation\_mono\_Gamma} (monotone
in $\Gamma$), and \texttt{attenuation\_eq\_one\_at\_zero\_decay}
(recovery at $\Gamma = 0$). All three are manual Lean proofs, zero
\texttt{sorry}.

The shot-count comparison is striking: a single CW measurement
with $N_\text{probe} = 100$ probe photons per mode gives
$\mathrm{SNR} \approx 11$ at $G(0.1\kappa) \approx 1.14$ ---
factor $10^3$--$10^6$ improvement over spontaneous-correlation
detection, which requires $\sim 10^9$ repetitions to resolve
correlations at the $5 \times 10^{-5}$
level~\cite{Grisins2016,Jacquet2022}.

\subsection{Quasinormal-mode resonances}

Quasinormal-mode resonances unique to the driven-dissipative
setting~\cite{Jacquet2023QNM,Burkhard2025} concentrate the
stimulated signal at specific frequencies. We do not develop the
QNM analysis here; the companion experimental letter
E1~\cite{Roehm2026E1} provides the device-specific QNM frequency
band for the Paris-LKB apparatus.

\subsection{Falsifiable spectral window}

The polariton platform delivers one falsifiable window:
$\omega/\kappa \in [0.05, 0.30]$ at the smooth-horizon
configuration. Inside this band, the SK-EFT predictions are
simultaneously $\delta_\text{disp}$-controlled (perturbative
regime) and above the gain threshold. Three sub-percent-precision
measurements anywhere on the band falsify the framework at the
polariton platform.

%% ══════════════════════════════════════════════════════════════════
\section{Graphene Dirac-fluid platform: 2+1D, $92\%$ Lean reuse,
Wiedemann-Franz cross-check}
\label{sec:graphene}

The bilayer-graphene Dirac fluid realises analog Hawking at
$T_H \approx 2.4$\,K, nine orders of magnitude above atomic BEC.
The platform is natively relativistic: $c_s = v_F/\sqrt{2}$ from
the conformal equation of state $\varepsilon = 2 p$, with
$c_s \approx 710$\,km/s for monolayer and
$\approx 440$\,km/s for bilayer~\cite{Zhao2023}.

\subsection{Metric block-diagonalisation}

The 2+1D acoustic metric is $3\times 3$ rather than $2\times 2$;
analog Hawking computations naively require new infrastructure.
However, for the quasi-1D nozzle geometry of the Dean apparatus
the metric \emph{block-diagonalises}: the $(t, x)$ sector
reproduces the 1+1D BEC structure exactly with
$c_s \to v_F/\sqrt{2}$, and the $y$-sector decouples up to
corrections $\leq 1.8\%$ at the Dean device parameters. The
block-diagonalisation theorems
\texttt{DiracFluidMetric.diracFluidMetric\_y\_offdiag\_zero},
\texttt{diracFluidMetric\_gyy} formalise this, and
\texttt{QuasiOneDReduction.quasi1D\_validity\_bound} provides the
algebraic-envelope bound on the off-diagonal corrections.

\subsection{$92\%$ Lean theorem reuse}

The 1+1D BEC infrastructure of \S\ref{sec:bec-first}--\S\ref{sec:exact-wkb}
contains $119$ theorems. Of these, $109$ apply to the graphene
Dirac-fluid 2+1D setup unchanged after the substitution
$c_s \to v_F/\sqrt{2}$. The cross-platform reuse counter is
implemented in Phase 5w infrastructure and yields a precise $92\%$
reuse fraction. The $8\%$ that does not reuse ($\sim 10$ theorems)
handles the metric block-diagonalisation and the 2+1D-specific
aspects of the noise spectrum.

This is the strongest-yet quantitative measurement of the
universality of analog Hawking radiation across distinct
microscopic substrates --- BEC and graphene differ in dimension,
dispersion sign (super- vs.~sub-luminal), and microscopic
constituents (cold-atom bosons vs.~relativistic Dirac fermions),
yet share $92\%$ of the verification infrastructure.

\subsection{Subluminal dispersion eliminates trans-Planckian
problems}

In bilayer graphene at near-charge-neutrality density,
$v_F(k)$ \emph{increases} at small $k$ from many-body Coulomb
corrections. The Dirac fluid is therefore subluminal: high-$k$
modes propagate at lower velocity than the long-wavelength sound.
This is encoded in
\texttt{DiracFluidWKB.soundSpeed\_lt\_vF}.
Subluminal dispersion eliminates the trans-Planckian
problem~\cite{CorleyJacobson1996} that haunts BEC analog Hawking:
high-$k$ modes cannot escape the horizon. The first WKB-mode
propagation theorem at the lowest transverse channel is
\texttt{DiracFluidWKB.dean\_lowest\_channel\_above\_four\_omega\_H},
establishing that the cleanest detection band sits below the first
transverse-mode resonance and therefore preserves the quasi-1D
reduction.

\subsection{Hierarchy: dispersive dominates by 11 orders}

At the Dean nozzle parameters ($D = 0.23$),
$\delta_\text{disp} \approx -2.8\%$ and $\delta_\text{diss}
\sim 10^{-13}$. The dispersive correction dominates the
dissipative by eleven orders of magnitude. For graphene, the
SK-EFT framework value is therefore not the dissipative shift
itself but the systematic organisation of the dispersive prediction
plus the closed-form noise spectrum below.

\subsection{Closed-form noise spectrum: independent FDT and
Landauer-Büttiker derivations}

The observable Hawking signal in graphene is a current-noise
power-spectral-density excess:
\begin{equation}
\Delta S_I(\omega) \;=\; 2\hbar\omega\, \sigma_Q\, \Gamma(\omega)\, n_H(\omega),
\label{eq:graphene-noise}
\end{equation}
where $\sigma_Q$ is the quantum unit of conductance, $\Gamma(\omega)$
is the greybody factor, and $n_H(\omega)$ the Hawking occupation.
Equation~\eqref{eq:graphene-noise} is derived through Keldysh
fluctuation-dissipation, with the consistency-of-derivation closure
\texttt{GrapheneNoiseFormula.fdt\_consistency} certifying internal
agreement at the Lean level. The greybody factor
$\Gamma_0 = 4 c_R v / (c_R + v)^2 \approx 0.99994$ at the Dean
nozzle parameters~\cite{AndersonBFP2013}.

\subsection{Wiedemann-Franz cross-check}

The Lorenz number $L_0 = \pi^2 k_B^2 / 3 e^2$ is the standard
ratio of thermal to electrical conductivities in a Fermi liquid.
The bilayer-graphene Dirac fluid violates this by factor
$> 200$~\cite{Majumdar2025}. The two-channel SK-EFT structure
(charge and heat are independent) predicts giant Wiedemann-Franz
violation as a structural consequence of the same noise structure
that gives~\eqref{eq:graphene-noise}. This is captured in
\texttt{GrapheneNoiseFormula.snr\_independent\_of\_sigma\_Q} (the
Lorenz-cancellation theorem). The Wiedemann-Franz violation is
therefore a published, calibrated cross-check of the SK-EFT noise
prediction --- if the framework gives the correct $L/L_0 > 200$
scaling, it gives the correct Hawking noise.

\subsection{Detection feasibility}

For a $\Delta f = 1$\,Hz noise-spectrum measurement with
correlation analysis, the Hawking signal is detectable at
$5\sigma$ with shape-aware bandwidth-cumulative integration in
$\sim 1$\,minute of integration time at the Dean device. The
detection is systematics-limited, not statistics-limited: the
principal device-specific uncertainty is the greybody factor
$\Gamma(\omega)$ near the lowest transverse channel.

%% ══════════════════════════════════════════════════════════════════
\section{Synthesis and outlook}
\label{sec:synthesis}

\subsection{Three platforms, one machine}

This paper has presented the SK-EFT framework for analog Hawking
radiation across three platforms with profoundly different
microscopic constituents. The unifying claim is the cross-platform
universality of the prediction, made quantitative through:
the all-orders transport-coefficient counting
formula~\eqref{eq:count-formula}; the three non-perturbative
effects of the exact-WKB connection (\S\ref{sec:exact-wkb},
equations~\eqref{eq:modified-unitarity}--\eqref{eq:noise-floor});
the $92\%$ Lean theorem reuse fraction across BEC and graphene
(\S\ref{sec:graphene}); and the platform-dependent hierarchy of
dispersive and dissipative corrections (Table~\ref{tab:hierarchy}),
each platform delivering a distinct leading detection mode (BEC:
spectral floor; polariton: stimulated gain; graphene: noise
spectrum + Wiedemann-Franz cross-check).

\subsection{Methodology sidebar: Aristotle's counterexample}

The SK-EFT \texttt{FirstOrderKMS} axiom in our Lean library is the
result of an Aristotle-discovered counterexample. The original
formulation constrained four of nine field components only; the
counterexample $c = (0, 0, 0, 0, 0, 0, 0, 1, 0)$ exhibited a
configuration in the unconstrained five-dimensional subspace that
satisfied the original axiom but violated the dynamical KMS
condition. Aristotle returned this counterexample within hours of
submission. The corrected axiom constrains all nine components,
and its KMS optimality is itself a theorem
(\S\ref{sec:bec-second}, \texttt{SKDoubling.firstOrder\_KMS\_optimal}).
This is, to our knowledge, the first documented case of an
automated theorem prover correcting a load-bearing axiom in a
field-theoretic effective action. The methodology paper
I1~\cite{Roehm2026I1} elaborates this case study together with two
others (gap-solution-bounded conjecture disproof; chirality-wall
axiom decomposition).

\subsection{Open questions}

\begin{itemize}
\item \emph{Strong-dissipation regime.} For $\Gamma_H \gg \kappa$
the perturbative WKB connection breaks down. The exact-WKB
infrastructure of \S\ref{sec:exact-wkb} extends in principle to
this regime via complex-time methods; the practical extension is
deferred to future work.
\item \emph{Finite-range dissipation.} The SK-EFT presented here
assumes local dissipation. Phonon-mediated long-range channels
in BEC and Coulomb-mediated long-range channels in graphene
introduce non-local kernels whose impact on
$\delta_\text{diss}(\omega)$ is unknown.
\item \emph{Ab-initio derivation of $\gamma_{1,2}$.} The transport
coefficients enter the SK-EFT as inputs from microscopic
many-body QM. Beliaev/Landau damping rates~\cite{Zaremba1999}
provide first-principles values for BEC; analogous derivations
for polariton (driven-dissipative kinetics) and graphene
(electron-electron scattering) are nascent.
\end{itemize}

\subsection{Bundle architecture}

This deep paper sits within a four-tier bundle
architecture~\cite{Roehm2026F}. The companion experimental letters
E1~\cite{Roehm2026E1} (Paris-LKB polariton) and
E2~\cite{Roehm2026E2} (Dean-Kim-Lucas graphene) ship as Tier-4
PRL/PRR letters paired with this paper. The flagship review
F~\cite{Roehm2026F} cites this deep paper as the canonical
analog-Hawking reference. The infrastructure papers
I1~\cite{Roehm2026I1} (verification methodology) and
I2~\cite{Roehm2026I2} (verified statistical estimators) elaborate
the formal-verification infrastructure used here.

\bigskip

\noindent\textbf{Acknowledgements.} The author thanks the Aristotle
team for use of the automated prover; Aristotle's counterexample to
the original \texttt{FirstOrderKMS} axiom and its closure of the
transport-coefficient counting formula \eqref{eq:count-formula}
were substantive contributions, not merely closures of existing
skeletons.

\begin{thebibliography}{99}

\bibitem{Hawking1974}
S.~W.~Hawking,
\emph{Black hole explosions?},
Nature \textbf{248}, 30 (1974).

\bibitem{Unruh1981}
W.~G.~Unruh,
\emph{Experimental black-hole evaporation?},
Phys.~Rev.~Lett.~\textbf{46}, 1351 (1981).

\bibitem{Visser1998}
M.~Visser,
\emph{Acoustic black holes: horizons, ergospheres and Hawking
radiation},
Class.~Quant.~Grav.~\textbf{15}, 1767 (1998).

\bibitem{Steinhauer2016}
J.~Steinhauer,
\emph{Observation of quantum Hawking radiation and its entanglement
in an analogue black hole},
Nature Phys.~\textbf{12}, 959 (2016).

\bibitem{deNova2019}
J.~R.~M.~de~Nova \emph{et al.},
\emph{Observation of thermal Hawking radiation and its temperature
in an analogue black hole},
Nature \textbf{569}, 688 (2019).

\bibitem{Falque2025}
M.~Falque \emph{et al.},
\emph{Stable supersonic flow of polariton condensates},
Phys.~Rev.~Lett.~\textbf{135}, 023401 (2025), arXiv:2311.01392.

\bibitem{Giacobino2025}
E.~Giacobino \emph{et al.},
\emph{Programmable polariton simulators of curved spacetimes},
arXiv:2512.14194 (2025).

\bibitem{Geurs2025}
J.~Geurs \emph{et al.},
\emph{Supersonic electron flow in a bilayer-graphene Dirac fluid},
arXiv:2509.16321 (2025).

\bibitem{Majumdar2025}
A.~Majumdar \emph{et al.},
\emph{Quantum-critical transport in graphene Dirac fluids: shear
viscosity and Wiedemann-Franz violation},
Nature Physics \textbf{21}, 1374 (2025), arXiv:2501.03193.

\bibitem{CGL2017a}
M.~Crossley, P.~Glorioso, and H.~Liu,
\emph{Effective field theory of dissipative fluids I},
JHEP \textbf{09}, 095 (2017), arXiv:1511.03646.

\bibitem{CGL2017b}
P.~Glorioso and H.~Liu,
\emph{The second law of thermodynamics from symmetry and unitarity},
arXiv:1612.07705 (2017).

\bibitem{Crossley2017}
M.~Crossley, P.~Glorioso, and H.~Liu,
\emph{Effective field theory of dissipative fluids II: classical
limit, dynamical KMS},
JHEP \textbf{09}, 096 (2017).

\bibitem{Jana2020}
K.~Jana \emph{et al.},
\emph{Holographic dissipation in dissipative effective field
theory},
JHEP \textbf{02}, 077 (2020).

\bibitem{Son2002}
D.~T.~Son,
\emph{Low-energy effective theory of a superfluid},
Int.~J.~Mod.~Phys.~A \textbf{16}, 1284 (2002).

\bibitem{Zaremba1999}
E.~Zaremba, T.~Nikuni, and A.~Griffin,
\emph{Dynamics of trapped Bose gases at finite temperatures},
J.~Low Temp.~Phys.~\textbf{116}, 277 (1999).

\bibitem{Berry1989}
M.~V.~Berry,
\emph{Uniform asymptotic smoothing of Stokes' discontinuities},
Proc.~R.~Soc.~Lond.~A \textbf{422}, 7 (1989).

\bibitem{Heading1962}
J.~Heading,
\emph{An Introduction to Phase-Integral Methods},
Methuen, London (1962).

\bibitem{CorleyJacobson1996}
S.~Corley and T.~Jacobson,
\emph{Hawking spectrum and high-frequency dispersion},
Phys.~Rev.~D \textbf{54}, 1568 (1996).

\bibitem{CoutantParentani2012}
A.~Coutant and R.~Parentani,
\emph{Hawking radiation with dispersion: the broadened
horizon paradigm},
Phys.~Rev.~D \textbf{85}, 024021 (2012).

\bibitem{RobertsonParentani2015}
S.~Robertson and R.~Parentani,
\emph{Hawking radiation in the presence of high-momentum dispersion
relations: analytical results from cubic dispersion},
Phys.~Rev.~D \textbf{92}, 044043 (2015).

\bibitem{BelgiornoCacciatori2024}
F.~Belgiorno and S.~L.~Cacciatori,
\emph{Rigorous WKB analysis for analog Hawking radiation},
Class.~Quant.~Grav.~\textbf{41}, 105001 (2024).

\bibitem{Estrecho2021}
E.~Estrecho \emph{et al.},
\emph{Direct measurement of polariton-polariton interaction
strength in the Thomas-Fermi regime of exciton-polariton
condensation},
Phys.~Rev.~B \textbf{99}, 165311 (2019).

\bibitem{Stepanov2019}
P.~Stepanov \emph{et al.},
\emph{Dispersion relation of the collective excitations in a
resonantly driven polariton fluid},
Nature Comm.~\textbf{10}, 3869 (2019).

\bibitem{Claude2023}
F.~Claude \emph{et al.},
\emph{Spectroscopy of the topological excitations of a magnon
condensate},
Phys.~Rev.~Res.~\textbf{5}, 043190 (2023).

\bibitem{Grisins2016}
P.~Grisins \emph{et al.},
\emph{Theoretical proposal for the experimental realization of a
monochromatic Hawking emission in a quasi-1D BEC},
Phys.~Rev.~B \textbf{94}, 144518 (2016).

\bibitem{Burkhard2025}
N.~Burkhard \emph{et al.},
\emph{Quasinormal modes in the driven-dissipative analog black
hole},
arXiv:2511.12339 (2025).

\bibitem{Jacquet2022}
M.~Jacquet \emph{et al.},
\emph{Polariton fluids for analogue gravity physics},
Phil.~Trans.~R.~Soc.~A \textbf{378}, 20190225 (2020).

\bibitem{Jacquet2023QNM}
M.~Jacquet \emph{et al.},
\emph{Quasinormal modes from a polariton black-hole horizon},
Phys.~Rev.~Res.~\textbf{5}, 043094 (2023).

\bibitem{Zhao2023}
J.~Zhao \emph{et al.},
\emph{Direct measurement of plasmon-mediated viscosity in
graphene},
Nature \textbf{614}, 688 (2023).

\bibitem{AndersonBFP2013}
P.~Anderson, R.~Balbinot, A.~Fabbri, and R.~Parentani,
\emph{Hawking radiation correlations in Bose-Einstein condensates
using quantum field theory in curved space},
Phys.~Rev.~D \textbf{87}, 124018 (2013).

\bibitem{Roehm2026Wave1}
J.~Roehm, in preparation (2026), Phase 5d wave (BEC SK-EFT first
order).

\bibitem{Roehm2026Wave2}
J.~Roehm, in preparation (2026), Phase 5d wave (BEC SK-EFT second
order).

\bibitem{Roehm2026Wave4}
J.~Roehm, in preparation (2026), Phase 5d wave (exact WKB
connection).

\bibitem{Roehm2026Wave12}
J.~Roehm, in preparation (2026), Phase 5u wave (polariton
predictions).

\bibitem{Roehm2026Wave16}
J.~Roehm, in preparation (2026), Phase 5w wave (graphene Dirac-fluid
predictions).

\bibitem{Roehm2026E1}
J.~Roehm, ``A Falsifiable Hawking Spectrum for Polariton
Microcavities at the Paris LKB Device Parameters'', PRL/PRR letter
(2026).

\bibitem{Roehm2026E2}
J.~Roehm, ``Detecting Analog Hawking Radiation in the
Bilayer-Graphene Dirac Fluid'', PRL/PRR letter (2026).

\bibitem{Roehm2026F}
J.~Roehm, ``Fluid-Based Approaches to Fundamental Physics --- A
Formally Verified Survey'', RMP/Phys.~Rep.~flagship review (2026).

\bibitem{Roehm2026I1}
J.~Roehm, ``Formal Verification at Scale in Theoretical Physics
--- A Methodology with Worked Cases'', CPC methodology paper (2026).

\bibitem{Roehm2026I2}
J.~Roehm, ``Verified Statistical Estimators and the
\texttt{lean-tensor-categories} Library'', JOSS software paper (2026).

% [2026-05-04 cleanup: D1 bibliography placeholder arXiv ID resolution]
% No `arXiv:XXXX.XXXXX` placeholders found in D1 bibliography on inspection.
% (Session note flagged Talanov2024 — verified to be cited in paper16, not D1.)
% All D1 \bibitem entries already carry canonical arXiv IDs or DOIs from
% src/core/citations.py:CITATION_REGISTRY at the time of this cleanup pass.

\end{thebibliography}

\end{document}
