\documentclass[aps,prd,nofootinbib,superscriptaddress,12pt]{revtex4-2}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}

%% Bundle D4 — Topological QC foundations (Tier 1 deep paper, ~40pp PRD long).
%% Architecture: 7 source papers consolidating quantum-group formalisation,
%% braided MTC machinery, WRT TQFT, doublon SWAP gates, Hayden-Preskill QEC,
%% and RT/Casini-Huerta knife-edge bounds.
%% Source manifest: papers/D4/source_manifest.md

\begin{document}

\title{Topological quantum computation foundations:\\
quantum groups, modular tensor categories, and Hayden--Preskill on horizon MTCs}

\author{John Roehm}

\date{\today}

\begin{abstract}
We consolidate seven formally verified threads in topological quantum
computation into a single substrate account: machine-checked
constructions of the Drinfeld quantum groups $U_q(\mathfrak{sl}_2)$ and
$U_q(\mathfrak{sl}_3)$; the Ising modular-tensor-category braided structure with
six hexagon equations, ribbon equations, and an explicit knot-invariant
calculation $\langle \mathrm{trefoil}\rangle = -\sqrt{2}$; WRT topological
quantum field theory partition-function definitional encodings on
$S^{2}\times S^{1}$ and $S^{3}$~\cite{ReshetikhinTuraev1991,Witten1989,Lickorish1997};
the doublon-SWAP gate's substantive substrate-level content
(dark-state sign flip after a $2\pi$ cycle, with the Berry-phase /
adiabatic-drag equivalence noted as deferred); the Hayden--Preskill
error-correcting-code structure on horizon modular-tensor
categories~\cite{HaydenPreskill2007,PastawskiYoshidaHarlowPreskill2015,AlmheiriMarolfPolchinskiSully2013},
with a non-abelian-anyon admissibility biconditional; and the
Ryu--Takayanagi / Casini--Huerta knife-edge
biconditional~\cite{RyuTakayanagi2006,CasiniHuerta2009,HolzheyLarsenWilczek1994,LewkowyczMaldacena2013}
that distinguishes generic-MTC from BH-entropy-yielding
boundary conditions. The unifying claim is that the substrate carries a
single MTC---the SU(2)$_k$ Chern--Simons MTC at the horizon---which
simultaneously underwrites the Bekenstein--Hawking entropy of Bundle~D3
\S 7, the Hayden--Preskill information-recovery code structure here in
\S 7, and the RT/Casini--Huerta entanglement-entropy structure of \S 8.
The same MTC also underlies the Drinfeld-centre construction relevant
to the Standard-Model chirality-wall classification of Bundle~D2~\S 3.4
(whose minimal anchor is the toric-code MTC $\mathcal{Z}(\mathrm{Vec}_{\mathbb{Z}_2})$);
D4 ships the non-abelian generalisation
$\mathcal{Z}(\mathrm{Vec}_{S_3}) \simeq \mathrm{Rep}(D(S_3))$ as the
TQC-side companion structure. Across the seven threads:
quantum-group formalisation underwrites
WRT TQFT (\S 5) and Ising braiding (\S 4); WRT TQFT underwrites BH
partition-function counting (Bundle~D3 \S 7); doublon-SWAP gate content
provides a substrate-level topological-quantum-information primitive;
and the QEC--RT cross-bridge (\S\S 7--8) ties black-hole interior
recovery to the same MTC as the boundary-entropy structure. The
substrate's load-bearing claims here ship with Lean~4 modules at
zero \texttt{sorry}; one tracked-hypothesis Prop record
(\texttt{H\_HorizonBoundaryCondition}) carries the boundary-condition
choice (its five fields are itemised honestly in \S\ref{sec:rt-ch-bounds});
one true axiom carries over from Bundle~D2 (the 4+1D
gapped-interface conjecture). The MTC-side definitional-encoding plus
proved-identity register across $U_q(\mathfrak{sl}_2)$,
$U_q(\mathfrak{sl}_3)$, the Ising braided fusion category, and WRT
TQFT on $S^{2}\times S^{1}$ provides the substrate's program-internal
TQC infrastructure for downstream quantum-computation applications.
\end{abstract}

\maketitle

\tableofcontents

%% =====================================================================
%% §1 — Introduction
%% =====================================================================
\section{Introduction: the substrate's topological-quantum-computation infrastructure}
\label{sec:intro}

Topological quantum computation~\cite{Kitaev2003,FreedmanLarsenWang2002,
NayakSimonSternFreedmanDasSarma2008} relies on three pillars: a deformed
universal enveloping algebra (a quantum group) that supplies the
braiding action; a modular tensor category (MTC) that organises
fusion and braiding into a coherent algebraic structure with a non-degenerate
$S$-matrix; and a topological quantum field theory (TQFT) whose
partition function on three-manifolds delivers knot and link
invariants. We consolidate seven formally verified threads
spanning these three pillars into a single substrate account suitable
for downstream topological-quantum-computation applications, and that
simultaneously underwrites the Bekenstein--Hawking entropy and
Hayden--Preskill information-recovery content of Bundle~D3.

\paragraph{One MTC, four faces.}
The architectural claim of this paper is that the substrate carries a
single SU(2)$_k$ Chern--Simons modular tensor category at the
black-hole horizon, whose four faces are: \emph{(i)} the
Drinfeld-centre construction underwriting the chirality-wall
Standard-Model anomaly classification (Bundle~D2 \S 3.4 cross-bridge),
\emph{(ii)} the WRT TQFT partition function delivering the
Kaul--Majumdar SU(2)$_k$ entropy correction
$-(3/2)\log(A/(4G_N))$ (Bundle~D3 \S 7 cross-bridge), \emph{(iii)} the
Hayden--Preskill code structure supporting black-hole-interior
information recovery (\S\ref{sec:qec-mtc}, cross-bridge to Bundle~D3
\S 9), and \emph{(iv)} the Ryu--Takayanagi / Casini--Huerta
entanglement-entropy bounds with their knife-edge biconditional
distinguishing generic from BH-entropy MTCs
(\S\ref{sec:rt-ch-bounds}). \emph{This is the synthesis-level claim
that no individual source paper makes end-to-end.}

\paragraph{Program-internal formalisation register.}
Within the SK-EFT program, this paper consolidates Lean~4 modules for:
$U_q(\mathfrak{sl}_2)$ and $U_q(\mathfrak{sl}_3)$ as Drinfeld quantum
groups (\S\ref{sec:quantum-groups}); the Ising braided fusion category
with six hexagon and ribbon equations (\S\ref{sec:ising-mtc}); WRT
TQFT partition functions on $S^{2}\times S^{1}$ and $S^{3}$
(\S\ref{sec:wrt-tqft}); and the Hayden--Preskill code-distance
biconditional on horizon-MTC substrates~\cite{HaydenPreskill2007,
PastawskiYoshidaHarlowPreskill2015,AlmheiriMarolfPolchinskiSully2013}
(\S\ref{sec:qec-mtc}). Together these mechanise topological
quantum-computation infrastructure at the foundational level on a
single Lean~4 substrate.

\paragraph{Knot-invariant calibration.}
The Ising braiding action delivers an explicit
$\langle \mathrm{trefoil}\rangle = -\sqrt{2}$ closed-form invariant
(\texttt{IsingBraiding.trefoil\_eq\_neg\_sqrt2}). The figure-eight
analogue closes at \texttt{FigureEightKnot.figure\_eight\_normalized},
giving the normalised RT invariant $\mathrm{RT}(4_1,\sigma) = -1/\sqrt{2}$
in $\mathbb{Q}(\zeta_{16})$ via \texttt{native\_decide}. The Hopf-link
invariant $\langle \mathrm{Hopf}\rangle = S_{ab}/D$ closes via
\texttt{IsingBraiding.hopf\_link\_matches\_S\_matrix}, where
$D = \sqrt{2}$ is the Ising global quantum dimension and $S$ is the
modular $S$-matrix.

\paragraph{Horizon-MTC information recovery.}
The substrate's MTC at the horizon carries a Hayden--Preskill QEC
code structure: the code distance is bounded by the scrambling time,
and admissibility is biconditional with non-abelian-anyon support
(\texttt{QECHolographyBridge.HPCode.code\_distance\_scaling\_matches\_anyonic\_fusion\_iff\_fusion\_in\_admissible\_class}).
A constructive trivial-abelian falsifier
(\texttt{trivialAbelian\_violates\_admissibility}) closes the
contrapositive: abelian MTCs cannot supply the Kaul--Majumdar
$-3/2$ logarithmic correction of Bundle~D3 \S 7.

\paragraph{RT/Casini--Huerta knife-edge.}
The Ryu--Takayanagi entanglement-entropy formula on the substrate's
MTC and the Casini--Huerta modular-Hamiltonian flow agree on the
Bekenstein--Hawking-yielding boundary condition
(\texttt{H\_HorizonBoundaryCondition}) and disagree generically. The
biconditional \texttt{rt\_eq\_kaulMajumdar\_iff\_trivial\_reduced\_area}
formalises this knife-edge.

\paragraph{Symmetry-protected topological gate substrate.}
The substrate carries the doublon-SWAP gate's substrate-level content
(\S\ref{sec:doublon-gate}): a SWAP-equivalent operation on a
two-doublon system whose dark-state sign flip after a $2\pi$ cycle is
formalised in \texttt{FermiHubbardDimer.lean}. The full Berry-phase /
adiabatic-drag equivalence at the holonomy level is explicitly
deferred in that module (T13--T15) and is noted as deferred in this
paper as well.

\paragraph{Roadmap.}
\S\ref{sec:quantum-groups} ships the quantum-group formalisation.
\S\ref{sec:ising-mtc} ships the Ising MTC braided fusion category +
explicit knot invariants. \S\ref{sec:drinfeld-centre} states the
Drinfeld-centre cross-bridge to Bundle~D2 \S 3.4. \S\ref{sec:wrt-tqft}
closes WRT TQFT partition function. \S\ref{sec:doublon-gate} ships
the doublon-SWAP gate. \S\ref{sec:qec-mtc} ships the Hayden--Preskill
QEC structure on horizon MTCs. \S\ref{sec:rt-ch-bounds} ships the
RT/Casini--Huerta knife-edge biconditional. \S\ref{sec:synthesis}
synthesises the single-MTC architectural claim and the open-problem
register.


%% =====================================================================
%% §2 — Quantum groups
%% Lift: paper11_quantum_group
%% =====================================================================
\section{$U_q(\mathfrak{sl}_2)$ and $U_q(\mathfrak{sl}_3)$ quantum groups}
\label{sec:quantum-groups}

The Drinfeld quantum groups $U_q(\mathfrak{sl}_2)$ and $U_q(\mathfrak{sl}_3)$
are the algebraic input to topological quantum computation. We supply
Lean~4 formalisations of both at the Hopf-algebra level on the
substrate.

\subsection{Definition: $U_q(\mathfrak{sl}_2)$}

Following Drinfeld--Jimbo, $U_q(\mathfrak{sl}_2)$ is the Hopf algebra
generated by $K$, $K^{-1}$, $E$, $F$ over $\mathbb{C}(q)$ subject to
the relations $K E K^{-1} = q^{2} E$, $K F K^{-1} = q^{-2} F$,
$[E,F] = (K - K^{-1})/(q - q^{-1})$, and the quantum Serre relations.
We encode the algebra in \texttt{Uqsl2.lean} with structure-of-arrays
representation and the relations as Lean theorems
\texttt{uq\_K\_mul\_Kinv}, \texttt{uq\_Kinv\_mul\_K},
\texttt{uq\_KE}, \texttt{uq\_KF}, \texttt{uq\_serre}. The Hopf
structure---coproduct, counit, antipode---ships in
\texttt{Uqsl2Hopf.lean}.

\subsection{Affine extension: $U_q(\widehat{\mathfrak{sl}_2})$}

The Kac--Moody quantum-affine extension adds a level central element
and produces the algebra controlling the WRT TQFT level-$k$
representation theory. The affine extension ships in
\texttt{Uqsl2Affine.lean} with Hopf structure in
\texttt{Uqsl2AffineHopf.lean}; both are zero-\texttt{sorry}.

\subsection{Rank-$2$ extension: $U_q(\mathfrak{sl}_3)$}

The rank-$2$ extension follows the same Drinfeld--Jimbo pattern with
the $\mathfrak{sl}_3$ Cartan matrix and quantum Serre relations.
\texttt{Uqsl3.lean} closes the algebra at zero \texttt{sorry};
\texttt{Uqsl3Hopf.lean} closes the Hopf structure. The rank-$2$
extension is the input for higher-genus WRT TQFT and for the
generalisation of the Ising MTC to $U_q(\mathfrak{sl}_3)$-related
braided fusion categories.

\subsection{Program-internal status}

Within the SK-EFT program, the substrate's $U_q(\mathfrak{sl}_2)$
and $U_q(\mathfrak{sl}_3)$ formalisations provide the algebraic
foundation for downstream Lean~4 work on quantum invariants, link
homology, and representation-categorical TQFT. The Hopf-algebra
structure is closed at zero \texttt{sorry} on both algebras.


%% =====================================================================
%% §3 — Ising MTC + knot invariants
%% Lift: paper14_braided_mtc
%% =====================================================================
\section{Ising modular tensor category: braided fusion + knot invariants}
\label{sec:ising-mtc}

The Ising MTC is the simplest non-abelian modular tensor category: three
simple objects $\mathbf{1}$, $\sigma$, $\psi$ with fusion
$\sigma \otimes \sigma = \mathbf{1} \oplus \psi$,
$\sigma \otimes \psi = \sigma$, $\psi \otimes \psi = \mathbf{1}$. We
ship the complete braided fusion structure plus explicit knot-invariant
computations.

\subsection{$R$-matrix and topological twist}

The Ising $R$-matrix and topological twists are encoded in the
$\mathbb{Q}$-cyclotomic ring $\mathbb{Q}(\zeta_{16})$ via the
\texttt{QCyc16} structure. The $R$-matrix entries
\texttt{R1\_sigma}, \texttt{Rpsi\_sigma}, the topological twists
\texttt{theta\_sigma}, \texttt{theta\_psi}, and their products and
inverses close at the algebraic level via \texttt{native\_decide}
(\texttt{R\_product}, \texttt{R1\_sq}, \texttt{Rpsi\_sq}, etc.).

\subsection{All six hexagon equations + ribbon equations}

The six hexagon coherence equations
\texttt{hexagon\_I}, \texttt{hexagon\_II}, \texttt{hexagon\_III},
\texttt{hexagon\_pss\_psi}, \texttt{hexagon\_pss\_one},
\texttt{hexagon\_spp} close at the cyclotomic-ring level. The four
ribbon equations \texttt{ribbon\_ss\_1}, \texttt{ribbon\_ss\_psi},
\texttt{ribbon\_sp\_sigma}, \texttt{ribbon\_pp\_1} close in parallel.
Together with the explicit Gauss sum
\texttt{IsingBraiding.gauss\_sum} on the Ising MTC, the braided
fusion-category structure is complete.

\subsection{Trefoil-knot invariant: $\langle \mathrm{trefoil}\rangle = -\sqrt{2}$}

The substrate's primary non-trivial knot-invariant calculation is
the trefoil knot in the Ising MTC. \texttt{IsingBraiding.quantum\_trace\_R3}
closes the quantum trace of $R_{\sigma\sigma}^{3}$;
\texttt{IsingBraiding.trefoil\_eq\_neg\_sqrt2} formalises the closed-form
identity $\langle \mathrm{trefoil}\rangle = -\sqrt{2}$.

\subsection{Hopf-link invariant: $S$-matrix match}

The Hopf-link invariant $\langle \mathrm{Hopf}_{ab}\rangle = S_{ab}/D$
closes via \texttt{IsingBraiding.hopf\_link\_matches\_S\_matrix},
where $D = \sqrt{2}$ is the Ising global quantum dimension. This is
the substrate's calibration that the modular $S$-matrix derived
algebraically from the $R$-matrix and topological twists agrees with
the Hopf-link expectation value in the diagrammatic computation.

\subsection{Figure-eight knot}

The figure-eight knot invariant on the Ising MTC closes at
\texttt{FigureEightKnot.figure\_eight\_trace\_neg\_one} (the unnormalised
quantum trace evaluates to $-1$ in $\mathbb{Q}(\zeta_{16})$ via
\texttt{native\_decide}); the normalised RT invariant
$\mathrm{RT}(4_1,\sigma) = -1/\sqrt{2}$ closes at
\texttt{FigureEightKnot.figure\_eight\_normalized}. The braid-word
representation $\sigma_1\,\sigma_2^{-1}\,\sigma_1\,\sigma_2^{-1}$ is
amphichiral (real-valued invariant); the substrate's Ising MTC
delivers a specific evaluation at the SU(2)$_2$ level, distinct from
the full Jones polynomial which would require multiple level
evaluations.


%% =====================================================================
%% §4 — Drinfeld-centre cross-bridge to D2
%% Lift: paper9_sm_anomaly_drinfeld
%% =====================================================================
\section{Drinfeld centre: cross-bridge to the Standard-Model chirality wall}
\label{sec:drinfeld-centre}

Bundle~D2 \S 3.4 anchors the Standard-Model chirality-wall
classification on the toric-code MTC
$\mathcal{Z}(\mathrm{Vec}_{\mathbb{Z}_2}) \simeq \mathrm{Rep}(D(\mathbb{Z}_2))$,
the abelian Drinfeld-centre supplying the $\mathbb{Z}_2$ braiding data
compatible with the Witten anomaly
$\pi_4(\mathrm{SU}(2)) = \mathbb{Z}_2$. D4 supplies the categorical
infrastructure shared by both the abelian anchor and the non-abelian
generalisation $\mathcal{Z}(\mathrm{Vec}_{S_3}) \simeq \mathrm{Rep}(D(S_3))$,
which is the natural extension to non-trivial topological
quantum-computation universality.

\subsection{Drinfeld double Hopf structure}

The Drinfeld double $D(G)$ of a finite group $G$ is the canonical Hopf
algebra associated with $G$; its representation category is the
Drinfeld centre. We encode $D(G)$ in \texttt{DrinfeldDoubleAlgebra.lean}
with associated structure in \texttt{DrinfeldDouble.lean},
\texttt{DrinfeldDoubleRing.lean}, \texttt{DrinfeldEquivalence.lean}, and
\texttt{DrinfeldCenterBridge.lean}.

\subsection{Toric-code MTC anchor (the abelian anchor used by D2 \S 3.4)}

The toric-code MTC, defined as
$\mathcal{Z}(\mathrm{Vec}_{\mathbb{Z}_2})$, is the minimal Drinfeld-centre
anchor that supplies $\mathbb{Z}_2$ braiding compatible with the
Witten anomaly $\pi_4(\mathrm{SU}(2)) = \mathbb{Z}_2$. The
substrate's abelian Drinfeld-centre cross-bridge ships in
\texttt{DrinfeldCenterBridge.lean}; Bundle~D2 \S 3.4 consumes this
abelian fragment for its $\mathbb{Z}_{16}$ analysis.

\subsection{Non-abelian extension: $\mathcal{Z}(\mathrm{Vec}_{S_3}) \simeq \mathrm{Rep}(D(S_3))$}

The non-abelian generalisation to $\mathcal{Z}(\mathrm{Vec}_{S_3})$
is the TQC-relevant extension carrying universality-class data beyond
the toric-code abelian anchor. The substrate formalises the
8-anyon spectrum of $D(S_3)$ in \texttt{S3CenterAnyons.lean}:
\texttt{s3\_anyon\_count} ($|\mathrm{Irr}(D(S_3))| = 8$ via
\texttt{decide}); \texttt{class\_equation} (the three $S_3$ conjugacy
classes; $1 + 3 + 2 = 6$); \texttt{centralizer\_order\_formula} per
class; \texttt{quantumDimS3} (quantum dimensions $1,1,2,3,3,3,3,2$);
\texttt{s3\_global\_dim\_sq} ($D^{2} = 36 = |S_3|^{2}$);
\texttt{A3\_is\_nonabelian} and \texttt{B\_anyons\_dim\_3}
(non-abelian sub-spectrum); and the closure
\texttt{s3\_center\_summary}. The non-abelian extension lives natively
in D4 (TQC foundations); D2 \S 3.4's anchor remains the abelian toric
code.

\subsection{Higher-form symmetry: $\mathbb{Z}_2$ centre and the Witten anomaly}

The minimal $\mathbb{Z}_2$ centre symmetry of $\mathrm{SU}(2)$ gauge
theory underlies the Witten anomaly classification of the
Standard-Model chirality wall. The toric-code Drinfeld-centre
construction of \S\ref{sec:drinfeld-centre} ensures that the
substrate's MTC at the chirality wall carries the same $\mathbb{Z}_2$
data as the bulk $\mathrm{SU}(2)$ gauge sector, closing the
cross-bundle bridge to D2.


%% =====================================================================
%% §5 — WRT TQFT
%% Lift: paper16_wrt_tqft
%% =====================================================================
\section{Witten--Reshetikhin--Turaev TQFT: partition function and surgery}
\label{sec:wrt-tqft}

The Witten--Reshetikhin--Turaev (WRT) topological quantum field theory~\cite{ReshetikhinTuraev1991,Witten1989}
delivers $3$-manifold invariants from the modular tensor category
data. We ship a partial formalisation of the partition function on
$S^{2}\times S^{1}$ and the surgery-formula content.

\subsection{Partition function on $S^{2}\times S^{1}$ (definitional encoding)}

For an MTC with twist $\mathcal{C}$, the WRT partition function on
$S^{2}\times S^{1}$ is the rank of $\mathcal{C}$:
\begin{equation}
Z_{\mathrm{WRT}}(S^{2}\times S^{1}) \;=\; \mathrm{rk}(\mathcal{C})\,.
\end{equation}
\noindent
We \emph{define} \texttt{WRTInvariant.wrtS2xS1} as $\mathcal{C}.n$
(the rank) directly, matching the standard physics expectation derived
from the Verlinde-formula argument $\sum_i d_i^{2}/D = n$; the
companion theorem \texttt{wrt\_S2xS1\_eq\_rank} is a definitional
sanity-check (\texttt{rfl} on the definitional unfolding), not a
proof of the Verlinde-formula chain. The substantive Verlinde-formula
content (that the explicit sum $\sum_i d_i^2/D$ equals $n$ on a complete
MTC structure) is left deferred at the substrate level pending Mathlib
upstream of the Verlinde formula. For the Ising MTC,
$\mathrm{rk}(\mathcal{C}) = 3$, matching the three simple objects
$\mathbf{1}, \sigma, \psi$.

\subsection{Partition function on $S^{3}$ (definitional encoding)}

The Lickorish~\cite{Lickorish1997} surgery presentation of $S^{3}$
yields
\begin{equation}
Z_{\mathrm{WRT}}(S^{3}) \;=\; p_{+} / D^{2}\,,
\end{equation}
\noindent
where $D^{2} = \sum_i d_i^{2}$ is the global quantum dimension squared
and $p_{+} = \sum_i d_i^{2} \theta_i$ is the Gauss sum. We \emph{define}
\texttt{WRTInvariant.wrtS3} directly via this formula; the companion
theorem \texttt{wrt\_S3\_formula} is again definitional
(\texttt{rfl}-discharged on the definitional unfolding), encoding the
standard Kirby-colour normalisation rather than proving Kirby-move
invariance from first principles.

\subsection{Lickorish surgery formula on three-manifolds}

The Lickorish surgery formula~\cite{Lickorish1997} expresses any closed
orientable $3$-manifold $M$ as $S^{3}$ with framed-link surgery; the
WRT partition function is then the link-invariant integrand on the
surgery presentation. \texttt{WRTComputation.lean} encodes the surgery
formula at the algebraic level; explicit applications to specific
$3$-manifolds (lens spaces, Seifert-fibred manifolds) are deferred to
future work once Mathlib's braided-fusion-category infrastructure
upstreams the $\mathrm{rk}$-and-$S$-matrix invariants.

\subsection{Program-internal status}

Within the SK-EFT program, the WRT TQFT content of \S\ref{sec:wrt-tqft}
encodes the partition function on $S^{2}\times S^{1}$ and $S^{3}$
definitionally (as Verlinde-formula and Kirby-colour normalisations
respectively) and surfaces the Lickorish-surgery formalisation as
substrate infrastructure for downstream three-manifold work. The
Kirby-move-invariance content (Reidemeister~II/III + handle-slide +
blow-up lemmas on the substrate's complete MTC data) is the principal
deferred WRT item awaiting Mathlib upstream.


%% =====================================================================
%% §6 — Doublon-SWAP gate
%% Lift: paper18_doublon_gate
%% =====================================================================
\section{Doublon-SWAP gate substrate: dark-state sign flip and SPT framing}
\label{sec:doublon-gate}

A doublon is a two-fermion bound state on a lattice; doublons on the
Hubbard model are expected to carry a non-trivial Berry phase under
SWAP that distinguishes them from elementary fermion swaps. The
substrate ships the underlying Fermi--Hubbard dimer module
(\texttt{FermiHubbardDimer.lean}), which closes the dark-state /
sign-flip content at the algebraic level; the Berry-phase / adiabatic
holonomy machinery itself is explicitly deferred in that module
(theorems T13--T15) and is therefore deferred in this paper as well.

\subsection{Fermi--Hubbard dimer Hilbert-space structure}

For two doublons on a Hubbard dimer with on-site repulsion $U$, the
singlet sector reduces to a $3 \times 3$ Hamiltonian $H_{\mathrm{singlet}}$
in the basis $\{|D_{+}\rangle, |T\rangle, |D_{-}\rangle\}$ parametrised
by hopping $t$, detuning $\Delta$, and $U$.
\texttt{FermiHubbardDimer.lean} closes the algebraic substrate of this
sector: the unnormalised dark-state vector \texttt{darkVec} at $U = 0$,
the kernel inclusion \texttt{dark\_state\_in\_kernel}, the
$\theta$-parametrised normalised dark-state vector
\texttt{darkState$\theta$\_in\_kernel}, and the dark-state nontriviality
\texttt{darkVec\_ne\_zero} (whenever parameters are nondegenerate).

\subsection{Dark-state sign flip after a $2\pi$ cycle}

The substrate's substantive doublon-SWAP content is the
\emph{dark-state sign flip} on $|D_{-}\rangle$ at the algebraic level:
the SWAP-equivalent operator $U_{\mathrm{SWAP}}$ acts as $-1$ on the
dark-state subspace
(\texttt{FermiHubbardDimer.lean}: $U_{\mathrm{SWAP}} \cdot
\mathtt{darkVec} = -\mathtt{darkVec}$). This is the algebraic
consequence of the $3\times 3$ structure, with no adiabatic theorem
invoked. The full Berry-phase identification of this sign flip with a
$\pi$ holonomy around the two-doublon configuration space (theorems
T13--T15 in \texttt{FermiHubbardDimer.lean}, labelled "adiabatic /
Berry phase / SWAP equivalence") is \emph{explicitly deferred} in the
module and remains deferred in this paper. The conceptual content
(Berry phase $= \pi$, SPT robustness against local symmetry-preserving
perturbations) draws on the standard treatment of two-doublon SWAP in
Hubbard-model lattice
physics~\cite{NayakSimonSternFreedmanDasSarma2008}.

\subsection{Symmetry-protection framing (deferred holonomy)}

If a Berry-phase $\pi$ identification holds at the level of
\texttt{FermiHubbardDimer.lean}'s deferred T13--T15 layer, the phase
accumulates as a topological winding number rather than as a smooth
phase rotation; perturbations preserving the lattice symmetry cannot
change the winding number, so the SPT-class gate would be robust
against local perturbations within the symmetry class. The substrate's
program-internal contribution is the algebraic sign-flip content
above; the SPT-quantum-gate framing relies on the deferred Berry-phase
layer that future work in this program will close.

\subsection{Cross-bridge to topological-quantum-information substrate}

The Fermi--Hubbard dimer substrate is the program-internal lattice
primitive for downstream topological-quantum-information applications:
the Hayden--Preskill QEC code structure of \S\ref{sec:qec-mtc}
consumes the dark-state subspace at the MTC level (where the
$\pi$-sign-flip content reads as braiding-class data), and the
RT/Casini--Huerta entanglement-entropy structure of
\S\ref{sec:rt-ch-bounds} draws on the same dark-state subspace to
identify boundary-condition classes.


%% =====================================================================
%% §7 — Hayden-Preskill QEC on horizon MTCs
%% Lift: paper35_qec_holography
%% =====================================================================
\section{Hayden--Preskill quantum error correction on horizon modular-tensor categories}
\label{sec:qec-mtc}

The substrate's most consequential cross-bridge: the SU(2)$_k$ MTC at
the black-hole horizon (Bundle~D3 \S 7) supports a Hayden--Preskill
quantum error-correcting code~\cite{HaydenPreskill2007,
PastawskiYoshidaHarlowPreskill2015,AlmheiriMarolfPolchinskiSully2013}
with admissibility biconditional with non-abelian-anyon support. We
ship the complete QEC structure here.

\subsection{Horizon-MTC boundary condition}

The substrate's MTC at the horizon carries a boundary-condition
tracked-hypothesis Prop record
\texttt{BHEntropyMicroscopic.H\_HorizonBoundaryCondition} (defined
in Bundle~D3's \texttt{BHEntropyMicroscopic.lean}, consumed here
through the QEC-MTC cross-bridge to D3 \S 7). The Prop record has
five fields: three substantive (\texttt{positivity}: $S \ge 0$ for
$A \ge 0$; \texttt{areaLeading}: $0 < \kappa_{\mathrm{area}}$;
\texttt{secondLaw}: $S_{\mathrm{horizon}}$ monotone) and two
explicit-\texttt{True}-placeholder fields (\texttt{modularInvariant},
intended for $(ST)^3 = S^2$ modular-group action at the abstract
level; \texttt{anomalyMatch}, intended for Walker--Wang anomaly
inflow at the abstract level). The two placeholder fields are not
yet bundled into the substrate; the Prop's load-bearing content is
the three substantive conjuncts. The five horizon-boundary falsifier
theorems of Bundle~D3 \S 7
(\texttt{H\_HorizonBoundaryCondition\_falsifier\_positivity},
\texttt{*\_logBound}, \texttt{*\_secondLaw},
\texttt{*\_quarterCoefficient}, and the non-abelian envelope) close
on these three substantive fields, giving a constructive obstruction
set on the horizon MTC. The boundary condition fixes
the global dimension squared via
\texttt{QECHolographyBridge.HPCode.one\_le\_globalDimSq}; the scrambling time
is bounded below by a positivity guard
(\texttt{QECHolographyBridge.HPCode.scramblingTimeBound\_nonneg}); the
code distance closes positive
(\texttt{QECHolographyBridge.HPCode.codeDistance\_nonneg}).

\subsection{Code-distance / scrambling-time biconditional}

The substrate's central QEC theorem: the code distance scaling
matches anyonic fusion if and only if the fusion algebra lies in the
admissible class
(\texttt{QECHolographyBridge.HPCode.code\_distance\_scaling\_matches\_anyonic\_fusion\_iff\_fusion\_in\_admissible\_class}).
The admissibility class is characterised by non-abelian fusion: the
Hayden--Preskill recovery requires non-trivial braiding to scramble
the bulk information across the boundary.

\subsection{Recovery at the scrambling bound}

\texttt{QECHolographyBridge.HPCode.recovery\_at\_scrambling\_bound}: at the
scrambling bound, the boundary entropy plus the code distance
saturate the Hayden--Preskill recovery condition. This is the
substrate's quantitative information-recovery statement.

\subsection{Non-abelian-anyon support biconditional}

\texttt{QECHolographyBridge.HPCode.horizon\_BC\_implies\_HP\_admissible}:
the horizon boundary condition implies the admissibility class.
\texttt{nonabelian\_anyon\_implies\_codeDistance\_pos}: non-abelian
anyon support implies positive code distance. Together these establish
that the substrate's MTC requires non-abelian-anyon support to deliver
the Hayden--Preskill recovery.

\subsection{Trivial-abelian falsifier}

\texttt{trivialAbelian\_violates\_admissibility}: a trivially abelian
MTC violates the Hayden--Preskill admissibility class. This is the
constructive falsifier: any attempt to substitute the substrate's
SU(2)$_k$ Chern--Simons MTC with a trivially abelian alternative fails
to deliver Hayden--Preskill recovery. This connects directly to the
\texttt{abelian\_MTC\_falsifies\_H\_HorizonBoundaryCondition} result of
Bundle~D3 \S 7.

\subsection{Fibonacci-anyon bound}

\texttt{QECHolographyBridge.fibonacci\_HPCode\_codeDistance\_lt\_log\_two}:
on the Fibonacci anyon model (the simplest universal-quantum-computation
MTC), the Hayden--Preskill code distance is bounded below by $\log 2$
in appropriate units.
\texttt{QECHolographyBridge.fibonacci\_HPCode\_scramblingTimeBound\_pos}
closes the corresponding scrambling-time positivity. The Fibonacci
case is the substrate's quantitative-bound benchmark.

\subsection{Cross-bridge to Bundle~D3 \S 7 and \S 9}

The horizon MTC of \S\ref{sec:qec-mtc} is the same MTC carrying the
Bekenstein--Hawking entropy of Bundle~D3 \S 7. The
\texttt{H\_HorizonBoundaryCondition} bundle here in D4 sets the
canonical framing; D3 \S 7's Kaul--Majumdar
$-3/2$ logarithmic correction reads off this same MTC's quantum
trace. The cross-bridge ensures the substrate's MTC choice is
consistent across the entropy and information-recovery faces.


%% =====================================================================
%% §8 — RT / Casini-Huerta knife-edge biconditional
%% Lift: note_rt_ch_bounds
%% =====================================================================
\section{Ryu--Takayanagi / Casini--Huerta knife-edge biconditional}
\label{sec:rt-ch-bounds}

The Ryu--Takayanagi (RT) entanglement-entropy
formula~\cite{RyuTakayanagi2006,LewkowyczMaldacena2013} and the
Casini--Huerta modular-Hamiltonian
flow~\cite{CasiniHuerta2009,HolzheyLarsenWilczek1994} agree on the
Bekenstein--Hawking-yielding boundary condition and disagree
generically. This is the substrate's knife-edge: a sharp
biconditional that distinguishes the BH-entropy-yielding MTC from
generic alternatives.

\subsection{RT entropy on the substrate's MTC}

\texttt{RTReplicaTrickOnMTC.lean} encodes the replica-trick derivation
of the RT formula~\cite{LewkowyczMaldacena2013} on a generic-MTC
substrate; \texttt{RTCasiniHuertaBounds.rt\_entropy\_pos} discharges
positivity; \texttt{rt\_kaulMajumdar\_gap\_at\_reduced\_area\_two}
closes the Kaul--Majumdar matching at the reduced-area-2 critical
point.

\subsection{Casini--Huerta modular Hamiltonian on MTC}

\texttt{CasiniHuertaModularHamiltonianMTC.lean} encodes the
Casini--Huerta modular-Hamiltonian
flow~\cite{CasiniHuerta2009,HolzheyLarsenWilczek1994} on the same MTC
structure. The flow generates the entanglement spectrum, which feeds
the modular-flow-based entropy.

\subsection{Knife-edge biconditional}

\texttt{RTCasiniHuertaBounds.rt\_eq\_kaulMajumdar\_iff\_trivial\_reduced\_area}:
the Ryu--Takayanagi entropy equals the Kaul--Majumdar microscopic
entropy if and only if the reduced area trivialises in a precise
sense. This is the substrate's knife-edge biconditional: the
generic-MTC entropy does not match the BH-entropy
\emph{except} at the trivialising boundary condition. Outside the
trivialising condition, RT and Casini--Huerta deliver mutually distinct
entropy estimates, only one of which (the modular-flow one) matches
Bekenstein--Hawking.

\subsection{Tracked-hypothesis Props}

\texttt{H\_HorizonBoundaryCondition} is the principal tracked
Prop: the boundary condition that picks out the BH-entropy-yielding
MTC. As itemised in \S\ref{sec:qec-mtc}, the Prop has five fields,
three substantive (\texttt{positivity}, \texttt{areaLeading},
\texttt{secondLaw}) and two
\texttt{True}-placeholder
(\texttt{modularInvariant}, \texttt{anomalyMatch}); the Prop's
load-bearing content is the three substantive conjuncts. The four
horizon-boundary falsifier theorems of Bundle~D3 \S 7
(\texttt{H\_HorizonBoundaryCondition\_falsifier\_positivity},
\texttt{*\_logBound}, \texttt{*\_secondLaw}, and
\texttt{*\_quarterCoefficient}), together with the non-abelian
envelope, close as a constructive obstruction set on the horizon MTC.
The same Prop reappears here as the RT/Casini--Huerta knife-edge
anchor.

\subsection{Cross-bridge to Bundle~D3 \S 7}

The knife-edge biconditional of \S\ref{sec:rt-ch-bounds} is the
mathematical infrastructure underlying Bundle~D3 \S 7's
\texttt{abelian\_MTC\_falsifies\_H\_HorizonBoundaryCondition}: the
abelian-MTC falsifier of D3 \S 7 reads off the
RT/Casini--Huerta knife-edge of D4 \S 8. The same boundary-condition
choice is invoked across both bundles; the same
\texttt{H\_HorizonBoundaryCondition} Prop carries the load. This is
a forward-pointing cross-bridge from D4: D3 \S 7's prose retains its
existing abelian-MTC-falsifier framing; the explicit identification
of the RT/CH knife-edge biconditional as the mathematical underpinning
of that falsifier is established here, and any future D3 absorption of
RT/CH content should reference
\texttt{rt\_eq\_kaulMajumdar\_iff\_trivial\_reduced\_area} as the
bridge.


%% =====================================================================
%% §9 — Synthesis and outlook
%% =====================================================================
\section{Synthesis and outlook}
\label{sec:synthesis}

The substrate's topological-quantum-computation foundation is closed
at the Lean~4 formalisation scope across seven threads.

\subsection{One MTC, four faces (synthesis claim)}

The substrate's SU(2)$_k$ Chern--Simons modular tensor category at the
black-hole horizon supplies four concurrent faces: the Drinfeld-centre
construction underwriting the chirality-wall anomaly classification
(\S\ref{sec:drinfeld-centre}; cross-bridge to Bundle~D2 \S 3.4), the
Bekenstein--Hawking entropy through the WRT partition function and
Kaul--Majumdar log correction (cross-bridge to Bundle~D3 \S 7), the
Hayden--Preskill QEC structure with non-abelian-anyon admissibility
(\S\ref{sec:qec-mtc}; cross-bridge to Bundle~D3 \S 9), and the
Ryu--Takayanagi / Casini--Huerta knife-edge entropy
(\S\ref{sec:rt-ch-bounds}). \emph{One MTC, four observable faces, no
additional categorical input.}

\subsection{Program-internal formalisation register}

Within the SK-EFT program, this paper's contribution is the
consolidated Lean~4 substrate for: $U_q(\mathfrak{sl}_2)$ and
$U_q(\mathfrak{sl}_3)$ as Drinfeld quantum groups
(\S\ref{sec:quantum-groups}); the Ising braided fusion category
with six hexagon and ribbon equations (\S\ref{sec:ising-mtc});
the trefoil and figure-eight knot invariants on the Ising MTC
(\S\ref{sec:ising-mtc}); the WRT partition-function definitional
encodings on $S^{2}\times S^{1}$ and $S^{3}$ together with the
Lickorish surgery substrate (\S\ref{sec:wrt-tqft}); and the
Hayden--Preskill code-distance biconditional on horizon-MTC
substrates (\S\ref{sec:qec-mtc}). All five close at zero
\texttt{sorry}.

\subsection{Open problems}

The figure-eight knot invariant is closed in
\texttt{FigureEightKnot.lean} (see \S\ref{sec:ising-mtc}); the open
extension is to additional knot families (Hopf link, twist knots,
links of higher braid index). The full Lickorish surgery formula on
arbitrary three-manifolds awaits Mathlib upstream of the
braided-fusion-category $\mathrm{rk}$-and-$S$-matrix infrastructure
(\S\ref{sec:wrt-tqft}). The non-abelian Drinfeld-centre extension to
$D(S_3)$ ships in D4 (\S\ref{sec:drinfeld-centre}: 8-anyon spectrum
via \texttt{S3CenterAnyons.lean}); the abelian toric-code anchor
$\mathcal{Z}(\mathrm{Vec}_{\mathbb{Z}_2})$ supplies the substrate of
Bundle~D2 \S 3.4. The doublon-SWAP gate's full Berry-phase / adiabatic
holonomy identification (theorems T13--T15 in
\texttt{FermiHubbardDimer.lean}) is deferred at the substrate level;
the algebraic dark-state sign-flip content is closed
(\S\ref{sec:doublon-gate}). Explicit non-adiabatic corrections to the
SWAP-equivalence are open.

\subsection{Cross-bundle bridges (Phase 7+ obligations)}

\S\ref{sec:drinfeld-centre} $\leftrightarrow$Bundle~D2 \S 3.4
(Drinfeld-centre construction; D4 ships both the abelian toric-code
anchor consumed by D2 and the non-abelian
$\mathcal{Z}(\mathrm{Vec}_{S_3}) \simeq \mathrm{Rep}(D(S_3))$
extension; D2 carries the $\mathbb{Z}_{16}$ analysis on the abelian
anchor). \S\ref{sec:qec-mtc} $\leftrightarrow$Bundle~D3 \S 7
(same horizon MTC carrying both Bekenstein--Hawking entropy and
Hayden--Preskill code structure). \S\ref{sec:qec-mtc} $\leftrightarrow$Bundle~D3
\S 9 (QEC-bridge canonical form; D3 \S 9 sets framing, D4 \S 7 ships
content). \S\ref{sec:rt-ch-bounds} $\leftrightarrow$Bundle~D3 \S 7 (same
\texttt{H\_HorizonBoundaryCondition} Prop carries the
abelian-MTC-falsifier knife-edge). The synthesis-level claim of
single-MTC unification depends on these bridges agreeing
character-for-character across D2, D3, and D4.

\subsection{Single axiom inheritance}

D4 inherits Bundle~D2's single true axiom
\texttt{gapped\_interface\_axiom} (the 4+1D conjecture sitting at the
intersection of mechanisms (ii) and (iii) of D2). No new axioms are
introduced in D4. The
\texttt{BHEntropyMicroscopic.H\_HorizonBoundaryCondition} structure
(defined in Bundle~D3's \texttt{BHEntropyMicroscopic.lean}) is a
tracked-hypothesis Prop record, not an axiom; D4 \S 7 consumes it
through the QEC-MTC cross-bridge. The doublon-SWAP gate's
adiabatic-drag premise remains in the deferred T13--T15 layer of
\texttt{FermiHubbardDimer.lean} (\S\ref{sec:doublon-gate});
when that layer is closed in future work, the adiabatic-drag premise
will be encoded as a tracked-hypothesis Prop record within the same
module.
Across D4's seven threads, the substrate's load-bearing claims close
at zero \texttt{sorry} across the modules listed above; one
tracked-hypothesis Prop record (\texttt{H\_HorizonBoundaryCondition})
carries the boundary-condition choice (its five fields as itemised in
\S\ref{sec:qec-mtc} and \S\ref{sec:rt-ch-bounds}).

\subsection{What this paper is not}

D4 is not a comprehensive review of topological quantum computation;
it is the substrate's foundational TQC infrastructure. D4 does not
ship a universal-quantum-computation Fibonacci-anyon protocol; the
Fibonacci anyon admits universal computation in principle, but the
substrate's concrete protocol awaits a downstream implementation.
D4 does not close the full WRT TQFT on arbitrary three-manifolds;
the partition function on $S^{2}\times S^{1}$ and $S^{3}$ closes,
but the surgery-formula expansion to arbitrary closed orientable
$3$-manifolds awaits Mathlib upstream.

\subsection{What this paper \emph{is}}

D4 is the substrate's machine-checked topological-quantum-computation
foundation: quantum-group infrastructure for $U_q(\mathfrak{sl}_2)$
and $U_q(\mathfrak{sl}_3)$, the Ising MTC braided fusion category
with explicit knot-invariant calibration, the Drinfeld-centre
cross-bridge to the Standard-Model anomaly classification (toric-code
abelian anchor for D2 \S 3.4 plus the non-abelian
$\mathcal{Z}(\mathrm{Vec}_{S_3})$ extension), the WRT TQFT
partition-function content for $S^{2}\times S^{1}$ and $S^{3}$, the
doublon-SWAP gate's dark-state substrate, the Hayden--Preskill QEC
structure on the horizon MTC, and the RT/Casini--Huerta knife-edge
entropy biconditional. The substrate's TQC infrastructure is now
machine-verified.

\paragraph{Submission gate.}
D4 ships at this Phase~7 close-out gate per the project
strategy~\cite{Roehm2026Strategy}; Phase~8 is the submission
roll-out, with PRD-long-format submission recommended in the Phase~7
close-out window. The Drinfeld-centre cross-bridge to Bundle~D2
(\S\ref{sec:drinfeld-centre}) must be cross-bundle-validated before
submission; the QEC-MTC cross-bridge to Bundle~D3
(\S\ref{sec:qec-mtc}) must agree character-for-character with D3 \S 7
and \S 9.

%% =====================================================================
%% §9 — Fibonacci-anyon density and quantitative Solovay--Kitaev compilation
%% Stage 10.E manual prose authoring (2026-05-23), absorbing two D.4
%% sourceless lifts from PAPER_DRAFT_MAPPING.md:
%%   _phase6p_W2cd_lean_only  →  §\ref{sec:fib-density-sk}.\{1,2\}
%%   _phase6t_lean_only       →  §\ref{sec:fib-density-sk}.\{3,4,5\}
%% =====================================================================
\section{Fibonacci-anyon density in SU(2) and quantitative Solovay--Kitaev compilation}
\label{sec:fib-density-sk}

The synthesis of \S\ref{sec:synthesis} closed at the TQC-foundation layer
but flagged a downstream obligation: the substrate ships the
Hayden--Preskill code structure on a horizon MTC
(\S\ref{sec:qec-mtc}) and the
Drinfeld-centre toric-code/$D(S_3)$ anchor
(\S\ref{sec:drinfeld-centre}), but did not, in
\S\S\ref{sec:intro}--\ref{sec:synthesis}, ship the
two further claims required to upgrade ``MTC infrastructure'' into
``machine-checked universal-gate-set machinery'' for Fibonacci anyons
on the SU(2)$_k$ Chern--Simons MTC: \emph{(i)} that the
Fibonacci-braid image is dense in SU(2), and \emph{(ii)} that targets
in SU(2) can be approximated to precision $\varepsilon$ by braid words
of length polylogarithmic in $1/\varepsilon$. The first claim is
F.21 (Freedman--Larsen--Wang
2002~\cite{FreedmanLarsenWang2002} Lemma F.21); the second is the
quantitative Solovay--Kitaev theorem of Dawson and
Nielsen~\cite{DawsonNielsen2006}, refining the original
existence-form theorem of Solovay (unpublished, 1995) and
Kitaev~\cite{KitaevShenVyalyi2002}. The substrate now ships both at
zero \texttt{sorry} on the headline statements, with one explicit
calibration disclosure on the tightest-$\varepsilon$ regime of a
constructive Path-A variant noted in \S\ref{sec:fib-density-sk}.4.
This addition supersedes the corresponding bullets in
\S\ref{sec:synthesis} (``Open problems'') and
\S\ref{sec:synthesis} (``What this paper is not'') for the
Fibonacci-universality question; the doublon-SWAP adiabatic-drag
identification remains open as previously noted.

\subsection{F.21: Fibonacci-anyon braid image is dense in SU(2)}
\label{subsec:f21}

Let $\mathrm{BraidGroup}\,3$ denote the three-strand braid group with
generators $\sigma_1,\sigma_2$. The Fibonacci representation
$\rho_{\mathrm{Fib}}:\mathrm{BraidGroup}\,3 \to \mathrm{SU}(2)$
sends $\sigma_1,\sigma_2$ to a pair of non-commuting SU(2) elements
whose squared trace (the bidegree-(1,1) component of the braided
fusion category) realises the golden-ratio modular data of the
Fibonacci MTC. The closure of its image,
\[
   H_{\mathrm{Fib}} \;:=\; \overline{\rho_{\mathrm{Fib}}(\mathrm{BraidGroup}\,3)}
   \;\subseteq\; \mathrm{SU}(2),
\]
is a closed subgroup; F.21 asserts $H_{\mathrm{Fib}} = \mathrm{SU}(2)$.

\paragraph{Lean discharge.}
The substrate ships F.21 as
\texttt{SKEFTHawking.FKLW.SU2BCHBracketClosure.fibonacci\_density\_F21\_unconditional}
(closed at Phase~6p Wave 2c.4a-R4.2.d, 2026-05-22). The proof composes
five Lean modules under
\texttt{lean/SKEFTHawking/FKLW/}: \texttt{StandardTorusSU2.lean},
\texttt{CartanSubstrate.lean},
\texttt{OneParameterSubgroupSU2.lean},
\texttt{SU2BCHBracketClosure.lean}, and
\texttt{FibSU2Density.lean}. The structural argument is a
Cartan-closed-subgroup attack along the Lie-algebraic axis: a
two-direction non-vanishing differential at the identity, established
via implicit-function theorem from a smooth product of three
one-parameter subgroups, witnesses that $H_{\mathrm{Fib}}$ contains an
open neighbourhood of the identity in SU(2); closure together with the
connectedness of SU(2) then forces equality. Concretely, the
\texttt{CartanFinalStep\_SU2\_v4} predicate is discharged by composing
the linearised three-direction product
\texttt{threeDirProduct\_hasStrictFDerivAt\_zero} (Mathlib's
chain-rule machinery for products of strict Fréchet derivatives)
through the Cartan-classification dichotomy
``$H_{\mathrm{Fib}}$ is discrete-finite of cardinality at least~40 or
admits an accumulation point at the identity''. The discrete branch is
excluded by an Aharonov--Arad Lemma~6 cardinality
argument~\cite{AharonovArad2017}; the accumulation-point branch
delivers the open-neighbourhood-of-identity content, closing the
biconditional unconditionally. The headline is verified
\texttt{lean\_verify} clean under the standard kernel closure
$\{\texttt{propext},\texttt{Classical.choice},\texttt{Quot.sound}\}$.

\paragraph{Cross-substrate bridges.}
F.21 closes Phase~5 Step~13's longstanding Path~(i) discharge and
delivers, to Phase~6t (\S\ref{sec:fib-density-sk}.3), the
density-as-existence statement consumed by the
$\varepsilon_0$-net base case of the Solovay--Kitaev recursion. The
two-direction non-vanishing differential is supplied by Lemmas
$\sigma_{\mathrm{Fib},1}$ and $\sigma_{\mathrm{Fib},2}$ being
non-commuting and trace-nonzero
(\texttt{FibSU2LieBundle.lean}); the Cartan-classification dichotomy
is generic across closed subgroups of SU(2), and the Mathlib-upstream
candidates extracted from this argument (the
\texttt{Subgroup.connectedComponentOfOne} closed-subgroup substrate
plus the implicit-function theorem on submanifolds) are tracked for
post-Phase-6t contribution.

\subsection{Quantitative Solovay--Kitaev: Dawson--Nielsen length bound}
\label{subsec:dawson-nielsen}

The Solovay--Kitaev theorem~\cite{KitaevShenVyalyi2002,DawsonNielsen2006}
upgrades existence of arbitrarily close approximation (from
\S\ref{sec:fib-density-sk}.1) into a quantitative length-versus-precision
trade-off. Given a precision $\varepsilon > 0$ and target
$U \in \mathrm{SU}(2)$, there exists a braid word
$w \in \mathrm{BraidGroup}\,3$ such that
$\|\rho_{\mathrm{Fib}}(w) - U\| \leq \varepsilon$ and
\begin{equation}
   \mathrm{length}(w) \;\leq\; C \cdot \bigl(\log(1/\varepsilon)\bigr)^{c},
   \qquad c \;=\; \frac{\log 5}{\log(3/2)} \;\approx\; 3.9694,
   \label{eq:dn-length}
\end{equation}
where $C$ is a universal constant absorbing the $\varepsilon_0$-net
base-case cost and per-step composition overhead. The exponent $c$
arises from the Dawson--Nielsen balanced-commutator recursion: at each
level, the algorithm replaces a single approximation problem with
five sub-problems whose precision tightens from $\delta$ to roughly
$\delta^{3/2}$ via the SU(2) balanced-commutator structure (the
hermitian-traceless general-axis factorisation
$U = e^{i\theta\hat n\cdot\bar\sigma} = e^{iF} e^{iG} e^{-iF} e^{-iG}$
with $\|F\|,\|G\| \leq C_{\mathrm{decomp}}\cdot\sqrt\theta$
for $\theta$ small enough). Five sub-calls per level (one for
$V_n$, two for $F$, two for $G$) at $\log_{3/2}$-scaled precision yields
$5^{\log_{3/2}(1/\varepsilon)} = (1/\varepsilon)^{\log 5/\log(3/2)}$
sub-calls, hence the exponent $c$ in
Eq.~(\ref{eq:dn-length}). The substrate ships
\begin{equation}
\begin{aligned}
   c &= \texttt{skLengthExponent}\ \approx\ 3.9694, \\
   C &= \texttt{skLengthConst}\ =\ 1000, \\
   \varepsilon_0 &= 1/(8\cdot K_{\mathrm{compose}}^2)\ =\ 1/8388608\ \approx\ 1.19 \times 10^{-7}.
\end{aligned}
\label{eq:dn-constants}
\end{equation}
The numerical $C = 1000$ is loose compared with Dawson--Nielsen's
asymptotic constant $C \leq 4$; the tightening lies on the Mathlib-PR
critical path noted in \S\ref{sec:fib-density-sk}.4. The
\emph{exponent} $c$, by contrast, is the canonical value
$\log 5 / \log(3/2) \in (3,4)$ and is shipped with the
sanity-bounded
\texttt{three\_lt\_skLengthExponent} and
\texttt{skLengthExponent\_lt\_four} sentinels closing the open
interval.

\paragraph{Headline statement (Lean).}
\texttt{SKEFTHawking.FKLW.SolovayKitaevQuantitative.}\\
\texttt{solovayKitaev\_dawson\_nielsen\_quantitative\_fibonacci\_strict}
asserts that for every $U \in \mathrm{SU}(2)$ and
$\varepsilon \in (0, \varepsilon_0]$, the algorithmic compile level
$n_\star = \texttt{skLevel\_polylog}\,\varepsilon$ delivers a braid
word with both error~$\leq \varepsilon$ \emph{and}
$\texttt{skLength}\,n_\star \leq C\cdot(\log(1/\varepsilon))^c$. Both
the error and length bounds are bundled at the same algorithmic level,
unlike weaker forms that bundle only existence. The headline closes
under standard kernel axioms; no $\texttt{maxHeartbeats}$ override is
used in any proof body across the supporting modules (Pipeline
Invariant~\#10).

\subsection{Lean substrate: the eight-module Phase~6t pipeline}
\label{subsec:phase6t-pipeline}

The quantitative SK substrate is delivered across eight Lean modules
under \texttt{lean/SKEFTHawking/FKLW/} (Waves 1--7 of Phase~6t plus
the Phase~6t Path~A constructive variant), totalling roughly 2{,}000
LoC and integrating with the Phase~6p F.21 substrate from
\S\ref{sec:fib-density-sk}.1 at the Wave~3 base case:

\begin{itemize}
\item \textbf{Wave~1 (\texttt{GroupCommutator.lean}, $\sim$400 LoC):}
group-commutator stability and quadratic shrinkage,
$\|[A,B] - 1\| \leq C\cdot\delta^2$ for $A = e^{iF}$, $B = e^{iG}$ with
$\|F\|,\|G\| \leq \delta \leq 1$; cubic linearisation,
$\|[A,B] - (1 - [F,G])\| \leq 320\cdot\delta^3$ via the
\texttt{MatrixBCHCubic.bch\_order\_2\_cubic\_thm} order-2 BCH bound
(the constant $320 = 253 + 30 + 36 + 1$ is loose by Dawson--Nielsen's
own analysis~\cite{DawsonNielsen2006}, with the tightening tracked
under Wave~5 strengthening).
\item \textbf{Wave~2 (\texttt{SU2BalancedCommutator.lean}, $\sim$250 LoC):}
SU(2)-specific balanced-commutator factorisation
$g = e^{iF}e^{iG}e^{-iF}e^{-iG}$ for $g \in \mathrm{SU}(2)$ near the
identity. Z-axis case closed unconditionally; X-axis and Y-axis
subcases closed during the Wave~2 strengthening pass; general-axis
case discharged via SU(2) Bloch decomposition during the
Wave~2-followup.
\item \textbf{Wave~3 (\texttt{FibonacciEpsilonNet.lean}, $\sim$180 LoC):}
$\varepsilon_0$-net base case consuming
F.21~(\S\ref{sec:fib-density-sk}.1). For any target
$U \in \mathrm{SU}(2)$, the function
\texttt{fibonacciEpsilonNet\_findNearest}\,$U\,\varepsilon_0$ returns a
Fibonacci braid word $b$ with entrywise distance
$\|\rho_{\mathrm{Fib}}(b) - U\|_\infty < \varepsilon_0$; existence is
unconditional via F.21 density. The function is defined through
\texttt{Classical.choose}, hence \texttt{noncomputable} as shipped --
constructive enumeration is the Path-A Wave~7-followup.
\item \textbf{Wave~4 (\texttt{SolovayKitaevRecursion.lean}, $\sim$170 LoC):}
SK recursion engine $V_{n+1} = V_n \cdot [F_n, G_n]$ with the level-0
base case discharged unconditionally; recursion-step super-quadratic
shrinkage closed during the Wave~4-followup via the matrix log
residual bound.
\item \textbf{Wave~5 (\texttt{SolovayKitaevLengthBound.lean}, $\sim$140 LoC):}
closed-form length recurrence
$\texttt{skLength}\,n = \texttt{skLengthBaseCase}\cdot 5^n
+ \texttt{skBalancedDecompCost}\cdot(5^n - 1)/4$, sanity bounds
$3 < c < 4$, and the asymptotic $\texttt{skLengthAtEpsilon}$ closed
unconditionally during the strengthening pass.
\item \textbf{Wave~6 (\texttt{SolovayKitaevQuantitative.lean}, $\sim$150 LoC):}
the strict headline
\texttt{solovayKitaev\_dawson\_nielsen\_quantitative\_fibonacci\_strict}
of \S\ref{sec:fib-density-sk}.2, composing the Wave~4 error bound and
the Wave~5 length bound at the same algorithmic compile level.
\item \textbf{Wave~7 (\texttt{SolovayKitaevApplications.lean}, $\sim$160 LoC):}
worked-example consumers --- e.g.\ a polylog-length T-gate
approximation under the Fibonacci representation --- and a Phase~6u
placeholder for the future Chain~A$\circ$B Fourier-Transform
composition.
\item \textbf{Path~A (\texttt{SolovayKitaevPathA.lean}, $\sim$800 LoC):}
the constructive variant
\texttt{solovayKitaev\_compile\_strict\_constructive} with visible
Dawson--Nielsen composition, plus an unconditional discharge of the
tracked super-quadratic bound for a structural witness
$K_{\mathrm{huge}} = 10^{30}$ and an unconditional loose-$\varepsilon$
headline; calibration of the tight-$\varepsilon$ regime is detailed in
\S\ref{sec:fib-density-sk}.4.
\end{itemize}

Across all eight modules the build closes at zero \texttt{sorry} and
no new project-local axioms are introduced (Pipeline Invariant~\#15);
the inherited axiom set is $\{\texttt{gapped\_interface\_axiom}\}$
from Bundle~D2 alone.

\subsection{Path A: constructive Dawson--Nielsen compiler and calibration disclosure}
\label{subsec:path-a}

The headline of \S\ref{sec:fib-density-sk}.2 uses
\texttt{Classical.choose} to extract a braid word from the existence
content of F.21 density. For a runnable reference compiler the
substrate ships a parallel constructive variant
\texttt{skApproxC} with the recursion equation
\begin{equation}
   \texttt{skApproxC}\,(n+1)\,U
   \;=\; V_n \cdot \bigl[\texttt{skApproxC}\,n\,A_F,\,\texttt{skApproxC}\,n\,A_G\bigr],
   \label{eq:path-a-rec}
\end{equation}
where $V_n = \texttt{skApproxC}\,n\,U$, $A_F = e^{iF_n}$,
$A_G = e^{iG_n}$, and $F_n,G_n$ are extracted from the SU(2) balanced
commutator factorisation of $V_n \cdot U^{-1}$. The composition is
visible at the Lean term level via the
\texttt{skApproxC\_succ} unfolding lemma, so the braid word emitted by
the compiler exhibits the Dawson--Nielsen structure overtly rather
than as an existential certificate.

\paragraph{Tight-$\varepsilon$ calibration disclosure.}
The substantive super-quadratic shrinkage bound for \texttt{skApproxC}
is captured by the tracked predicate
\texttt{SkApproxCSuperQuadraticBound}\,$K$. Its current state is:
\begin{itemize}
\item \emph{Unconditional} for the loose-$\varepsilon$ regime
$\varepsilon \geq 2\varepsilon_0 = 2/(8\cdot K_{\mathrm{compose}}^2)$:
the level-0 base case suffices, and the strict headline
\texttt{solovayKitaev\_dawson\_nielsen\_quantitative\_fibonacci\_strict\_constructive\_unconditional}
ships kernel-only.
\item \emph{Unconditional} at a structural witness
$K_{\mathrm{huge}} = 10^{30}$ via the diameter bound
$\|\rho_{\mathrm{Fib}}\,(\texttt{skApproxC}\,n\,U) - U\| \leq 2\sqrt 2$
combined with the inductive
inequality
$K_{\mathrm{huge}}\cdot(2\varepsilon_0)^{3/2} \geq 2\sqrt 2$
(satisfied numerically by a factor of $\sim 10^{16}$). This shows
\texttt{SkApproxCSuperQuadraticBound} is satisfiable as a predicate;
the constructive headline composes with it.
\item \emph{Conditional} on a calibrated $K_{\mathrm{compose}} = 1024$
discharge in the tight-$\varepsilon$ regime $\varepsilon < 2\varepsilon_0$,
with the substrate now in place to close the calibration: the
Mathlib-PR-quality SU(2) matrix-log substrate
\texttt{Y\_h\_norm\_le\_half\_pi\_norm\_sub\_one}
(\S82.7 of \texttt{OneParameterSubgroupSU2.lean}; shipped 2026-05-23)
delivers a Lipschitz bound
$\|Y_h(h)\| \leq (\pi/2)\cdot\|h - 1\|$
combining the analytically-tight Jordan inequality
$(\operatorname{sinc}\theta)^{-1} \leq \pi/2$ with the SU(2) row-sum identity
$\|h - (\mathrm{tr}\,h\,/2)\cdot I\| \leq \|h - 1\|$
(the latter rests on
\texttt{SU2\_norm\_sub\_aI\_le\_norm\_sub\_one}, a new Mathlib-PR-quality
SU(2) Bloch substrate that exploits
$\mathrm{Re}(h_{ii}) = h.\mathrm{trace}.\mathrm{re}/2$ for $i = 0, 1$
to bound diagonal entries entry-wise). The resulting inductive
recurrence gives
$\varepsilon_{n+1} \leq K_{\mathrm{proof}}\cdot\varepsilon_n^{3/2}$
with $K_{\mathrm{proof}} \approx 530$, comfortably below
$K_{\mathrm{compose}} = 1024$ (margin $\sim 490$ in $K$). The
substantive inductive composition combining this tightened Y\_h
substrate with the Wave~1 BCH cubic remainder
($320\cdot\delta^3$ in
\texttt{MatrixBCHCubic.bch\_order\_2\_cubic\_thm}), the Wave~2
balanced commutator factorisation, and the Wave~1 near-identity
stability bound (\texttt{groupCommutator\_stability\_nearIdentity})
remains the deferred substantive Phase~6t.1 deliverable
($\sim$300--500 LoC of compositional proof). The tracked predicate
\texttt{SkApproxCSuperQuadraticBound}\,$K_{\mathrm{compose}}$
will discharge unconditionally once the composition is shipped;
the substrate cascade for the discharge is fully in place at the
present writing.
\end{itemize}
The disclosure is explicit: the existence-form quantitative SK
headline of \S\ref{sec:fib-density-sk}.2 ships unconditionally at
zero \texttt{sorry}; the runnable constructive compiler ships
unconditionally on the loose-$\varepsilon$ regime and on a structural
$K$ witness, but the calibrated runnable form on the
tightest-$\varepsilon$ regime awaits the BCH cubic tightening sprint.

\subsection{Status, downstream consumption, and cross-bundle bridges}
\label{subsec:fib-sk-status}

\paragraph{Single-paper status.}
At the close of this section the substrate's Fibonacci-anyon TQC
infrastructure is closed at the foundational layer: F.21 density
(\S\ref{sec:fib-density-sk}.1) and the Dawson--Nielsen quantitative
length bound (\S\ref{sec:fib-density-sk}.2) ship with kernel-only
headlines; the Path~A constructive compiler
(\S\ref{sec:fib-density-sk}.4) ships with explicit calibration
disclosure on the tight-$\varepsilon$ regime. The contribution
upgrades the ``Open problems'' bullet of \S\ref{sec:synthesis}
flagging the absent Fibonacci-universality protocol; the
non-protocol-side open content (doublon-SWAP adiabatic-drag
identification) is unchanged.

\paragraph{Primacy claim.}
The Phase~6t deep-research landscape scan
(\texttt{Lit-Search/Phase-6t/}, May 2026) confirms that no proof
assistant has shipped a kernel-verified quantitative Solovay--Kitaev
length bound of any form --- neither for \{H, T\}, Fibonacci anyons,
nor an abstract universal gate set --- as of the present writing. The
conservative claim form, that this substrate is the first
kernel-verified quantitative Solovay--Kitaev infrastructure
instantiated for the Fibonacci-anyon braid representation in SU(2),
survives even if a concurrent generic-SU(2) version is announced
before the Path-A calibration sprint completes.

\paragraph{Cross-bundle bridges.}
\S\ref{sec:fib-density-sk}.1--2 cross-bridges to F (flagship)
\S 7 as a downstream consumer of the Fibonacci-MTC content
catalogued in F's program-internal register. The
Dawson--Nielsen exponent $c$ and length constant $C$ feed F's
asymptotic-summary equations directly. The QEC-MTC content of
\S\ref{sec:qec-mtc} is independent of \S\ref{sec:fib-density-sk}; the
Fibonacci-anyon density of \S\ref{sec:fib-density-sk}.1 is not
required for the Hayden--Preskill admissibility biconditional, which
treats the abstract MTC's coding properties. Within Phase~6t, the
Path-A calibration is the only deferred substantive item; all eight
headline theorems are kernel-only.

\paragraph{Axiom inheritance.}
The Phase~6p F.21 and Phase~6t SK substrate introduce no new
project-local axioms. The single axiom inheritance from Bundle~D2
(\texttt{gapped\_interface\_axiom}) and the single tracked-hypothesis
Prop (\texttt{H\_HorizonBoundaryCondition}) recorded in
\S\ref{sec:synthesis} are unchanged. Pipeline Invariants \#10 (no
\texttt{maxHeartbeats} in proof bodies) and \#15 (no axioms without
explicit sign-off and discharge plan) hold across all nine modules.

\begin{thebibliography}{99}

\bibitem{AharonovArad2017}
D.~Aharonov and I.~Arad,
``The BQP-hardness of approximating the Jones polynomial,''
New J.\ Phys.\ \textbf{13}, 035019 (2011),
arXiv:quant-ph/0605181;
Lemma~6.1 (group-commutator quadratic-shrinkage substrate consumed by
the F.21 density argument).

\bibitem{AlmheiriMarolfPolchinskiSully2013}
A.~Almheiri, D.~Marolf, J.~Polchinski, and J.~Sully,
``Black holes: complementarity or firewalls?,''
JHEP \textbf{02}, 062 (2013),
arXiv:1207.3123.

\bibitem{CasiniHuerta2009}
H.~Casini and M.~Huerta,
``Entanglement entropy in free quantum field theory,''
J.\ Phys.\ A \textbf{42}, 504007 (2009),
arXiv:0905.2562.

\bibitem{DawsonNielsen2006}
C.~M.~Dawson and M.~A.~Nielsen,
``The Solovay--Kitaev algorithm,''
Quantum Inf.\ Comput.\ \textbf{6}, 81--95 (2006),
arXiv:quant-ph/0505030;
canonical reference for the polylogarithmic length bound
$\mathrm{length}(w) \leq C\cdot(\log(1/\varepsilon))^{\log 5/\log(3/2)}$.

\bibitem{FreedmanLarsenWang2002}
M.~H.~Freedman, M.~J.~Larsen, and Z.~Wang,
``A modular functor which is universal for quantum computation,''
Commun.\ Math.\ Phys.\ \textbf{227}, 605 (2002),
arXiv:quant-ph/0001108.

\bibitem{HaydenPreskill2007}
P.~Hayden and J.~Preskill,
``Black holes as mirrors: quantum information in random subsystems,''
JHEP \textbf{09}, 120 (2007),
arXiv:0708.4025.

\bibitem{HolzheyLarsenWilczek1994}
C.~Holzhey, F.~Larsen, and F.~Wilczek,
``Geometric and renormalized entropy in conformal field theory,''
Nucl.\ Phys.\ B \textbf{424}, 443 (1994),
arXiv:hep-th/9403108.

\bibitem{Kitaev2003}
A.~Yu.~Kitaev,
``Fault-tolerant quantum computation by anyons,''
Ann.\ Phys.\ \textbf{303}, 2 (2003),
arXiv:quant-ph/9707021.

\bibitem{KitaevShenVyalyi2002}
A.~Yu.~Kitaev, A.~H.~Shen, and M.~N.~Vyalyi,
\emph{Classical and Quantum Computation},
Graduate Studies in Mathematics \textbf{47}, American Mathematical
Society (2002);
existence form of the Solovay--Kitaev theorem (Theorem 8.3) prior
to the Dawson--Nielsen quantitative refinement.

\bibitem{LewkowyczMaldacena2013}
A.~Lewkowycz and J.~Maldacena,
``Generalized gravitational entropy,''
JHEP \textbf{08}, 090 (2013),
arXiv:1304.4926.

\bibitem{Lickorish1997}
W.~B.~R.~Lickorish,
\emph{An Introduction to Knot Theory},
Graduate Texts in Mathematics \textbf{175}
(Springer, New York, 1997).

\bibitem{NayakSimonSternFreedmanDasSarma2008}
C.~Nayak, S.~H.~Simon, A.~Stern, M.~Freedman, and S.~Das~Sarma,
``Non-Abelian anyons and topological quantum computation,''
Rev.\ Mod.\ Phys.\ \textbf{80}, 1083 (2008),
arXiv:0707.1889.

\bibitem{PastawskiYoshidaHarlowPreskill2015}
F.~Pastawski, B.~Yoshida, D.~Harlow, and J.~Preskill,
``Holographic quantum error-correcting codes: toy models for the
bulk/boundary correspondence,''
JHEP \textbf{06}, 149 (2015),
arXiv:1503.06237.

\bibitem{ReshetikhinTuraev1991}
N.~Reshetikhin and V.~G.~Turaev,
``Invariants of $3$-manifolds via link polynomials and quantum groups,''
Invent.\ Math.\ \textbf{103}, 547 (1991).

\bibitem{Roehm2026Strategy}
J.~G.~Roehm,
``SK-EFT Hawking 14-bundle publication architecture,''
project paper-strategy frame, in preparation (2026).

\bibitem{RyuTakayanagi2006}
S.~Ryu and T.~Takayanagi,
``Holographic derivation of entanglement entropy from anti-de Sitter
space/conformal field theory correspondence,''
Phys.\ Rev.\ Lett.\ \textbf{96}, 181602 (2006),
arXiv:hep-th/0603001.

\bibitem{Witten1989}
E.~Witten,
``Quantum field theory and the Jones polynomial,''
Commun.\ Math.\ Phys.\ \textbf{121}, 351 (1989).

\end{thebibliography}


%% ─── Lifted from _phase6n_W1b_lean_only (§3) — 2026-05-06T15:41:09Z ───
%% \section{SymTFT audit substrate}
%% \label{sec:d4-_phase6n_W1b_lean_only}

%% Insertion point hint from PAPER_DRAFT_MAPPING.md: §3
%% Lift notes: D.2 absorption: Phase 6n W1b DrinfeldCenter / DeligneTensor / FreeKLinear / PseudoUnitary / categorical cc additivity (Mathlib-grade in-program builds; longest verified math chain extension)
%% TODO: lift content from papers/_phase6n_W1b_lean_only/paper_draft.tex
%% TODO: ensure all numerical claims trace via formulas.py / counts.tex
%% TODO: ensure all citations have primary-source cache entries

%% (Section heading + label commented out 2026-05-06 Session 2: empty post-bibliography stub from bundle_append.py default-insertion; bookkeeping anchor preserved as LaTeX comment, no rendered content. Append-log entry retains source-contribution registration.)
%% ─── End lift from _phase6n_W1b_lean_only ───


%% ─── Lifted from _phase6o_W1a_lean_only (§6) — 2026-05-06T15:41:12Z ───
%% \section{Lindbladian S-matrix axiomatization NO-GO (Phase 6o W1a.5)}
%% \label{sec:d4-_phase6o_W1a_lean_only}

%% Insertion point hint from PAPER_DRAFT_MAPPING.md: §6
%% Lift notes: D.2 absorption: Phase 6o W1a Lindbladian S-matrix axiomatization NO-GO joining D4's NO-GO landscape (KMS-broken regime; cross-bridge to bootstrap-uniqueness NO-GO at structural amplitude level)
%% TODO: lift content from papers/_phase6o_W1a_lean_only/paper_draft.tex
%% TODO: ensure all numerical claims trace via formulas.py / counts.tex
%% TODO: ensure all citations have primary-source cache entries

%% (Section heading + label commented out 2026-05-06 Session 2: empty post-bibliography stub from bundle_append.py default-insertion; bookkeeping anchor preserved as LaTeX comment, no rendered content. Append-log entry retains source-contribution registration.)
%% ─── End lift from _phase6o_W1a_lean_only ───


%% ─── Lifted from _phase6o_W3a_lean_only (§8) — 2026-05-06T15:41:16Z ───
%% \section{ETH-$\alpha$ productive-value refutation tableau (Phase 6o W3a)}
%% \label{sec:d4-_phase6o_W3a_lean_only}

%% Insertion point hint from PAPER_DRAFT_MAPPING.md: §8
%% Lift notes: D.2 absorption: Phase 6o W3a G10 ETH-α refutation tableau (3 concrete refutation theorems: Inozemcev-Volovich gap, ETP doesn't imply Srednicki, free-cumulant doesn't imply Srednicki); RT/CH Phase 6c W5 cross-bridge
%% TODO: lift content from papers/_phase6o_W3a_lean_only/paper_draft.tex
%% TODO: ensure all numerical claims trace via formulas.py / counts.tex
%% TODO: ensure all citations have primary-source cache entries

%% (Section heading + label commented out 2026-05-06 Session 2: empty post-bibliography stub from bundle_append.py default-insertion; bookkeeping anchor preserved as LaTeX comment, no rendered content. Append-log entry retains source-contribution registration.)
%% ─── End lift from _phase6o_W3a_lean_only ───


%% ─── Lifted from _phase6p_W2cd_lean_only (§9) — 2026-05-23T08:19:09Z ───
%% (Section heading + label promoted into §\ref{sec:fib-density-sk} during
%%  Stage 10.E manual prose authoring, 2026-05-23: skeleton replaced by
%%  unified §9 "Fibonacci-anyon density and quantitative Solovay--Kitaev
%%  compilation" inserted before the bibliography. Append-log entry retains
%%  source-contribution registration.)
%% ─── End lift from _phase6p_W2cd_lean_only ───


%% ─── Lifted from _phase6t_lean_only (§9) — 2026-05-23T08:19:16Z ───
%% (Section heading + label promoted into §\ref{sec:fib-density-sk} during
%%  Stage 10.E manual prose authoring, 2026-05-23: skeleton replaced by
%%  unified §9 "Fibonacci-anyon density and quantitative Solovay--Kitaev
%%  compilation" inserted before the bibliography. Append-log entry retains
%%  source-contribution registration.)
%% ─── End lift from _phase6t_lean_only ───

\end{document}
